### 2017-11-16 ESSAN/article.tex
Draft

\documentclass{esannV2}
\usepackage[dvips]{graphicx}
\usepackage[latin1]{inputenc}
\usepackage{amssymb,amsmath,array}
\graphicspath{{images/}}

%***********************************************************************
% !!!! IMPORTANT NOTICE ON TEXT MARGINS !!!!!
%***********************************************************************
%
% Please avoid using DVI2PDF or PS2PDF converters: some undesired
% shifting/scaling may occur when using these programs
% It is strongly recommended to use the DVIPS converters, and to submit
% PS file. You may submit a PDF file if and only if you use ADOBE ACROBAT
% to convert your PS file to PDF.
%
% Check that you have set the paper size to A4 (and NOT to letter) in your
% dvi2ps converter, in Adobe Acrobat if you use it, and in any printer driver
% that you could use.  You also have to disable the 'scale to fit paper' option
% of your printer driver.
%
% In any case, please check carefully that the final size of the top and
% bottom margins is 5.2 cm and of the left and right margins is 4.4 cm.
% It is your responsibility to verify this important requirement.  If these margin requirements and not fulfilled at the end of your file generation process, please use the following commands to correct them.  Otherwise, please do not modify these commands.
%
\voffset 0 cm \hoffset 0 cm \addtolength{\textwidth}{0cm}
\addtolength{\textheight}{0cm}\addtolength{\leftmargin}{0cm}

%***********************************************************************
% !!!! USE OF THE esannV2 LaTeX STYLE FILE !!!!!
%***********************************************************************
%
% Some commands are inserted in the following .tex example file.  Therefore to
% set up your ESANN submission, please use this file and modify it to insert
% your text, rather than staring from a blank .tex file.  In this way, you will
% have the commands inserted in the right place.

\begin{document}
%style file for ESANN manuscripts
\title{LaTeX style file for ESANN manuscripts}

%***********************************************************************
% AUTHORS INFORMATION AREA
%***********************************************************************
\author{Ram\'on H. Mart\'inez-Mayorquin $^1$, Anders Lansner and $^2$ Pawel Herman $^3$
%
% Optional short acknowledgment: remove next line if non-needed
% \thanks{This is an optional funding source acknowledgement.}
%
% DO NOT MODIFY THE FOLLOWING '\vspace' ARGUMENT
\vspace{.3cm}\\
%
% Addresses and institutions (remove "1- " in case of a single institution)
1- KTH - Dept of First Author \\
Address of First Author's school - Country of First Author's
school
%
% Remove the next three lines in case of a single institution
\vspace{.1cm}\\
2- KTH - Dept of Second Author \\
Address of Second Author's school - Country of Second Author's school 
%
\vspace{.1cm}\\
3 School of Second Author - Dept of Second Author \\
Address of Second Author's school - Country of Second Author's school\\ 
}

%***********************************************************************
% END OF AUTHORS INFORMATION AREA
%***********************************************************************

\maketitle

\begin{abstract}
Type your 100 words abstract here. Please do not modify the style
of the paper. In particular, keep the text offsets to zero when
possible (see above in this `ESANNV2.tex' sample file). You may
\emph{slightly} modify it when necessary, but strictly respecting
the margin requirements is mandatory (see the instructions to
authors for more details).
\end{abstract}

\section{Introduction}

This is a sample file. Please use this file to correctly typeset a
submission to the ESANN conference. The associated pdf file will
help you to have an idea of what your paper should look like.

\subsection{Neural Network}

\begin{align}
\tau_m \dfrac{dx}{dt} &= \Phi(wx - \theta) \\
\tau_z \dfrac{dz}{dt} &= x - z
\end{align}

\begin{figure}[h!]
\centering
\includegraphics[scale=2.0]{diagram.eps}
\caption{ESANN 2005: Announcement and call for
papers.}\label{Fig:MV}
\end{figure}

Please avoid using DVI2PDF or PS2PDF converters: some undesired
shifting/scaling may occur when using these programs
It is strongly recommended to use the DVIPS converters, and to submit
PS file. You may submit a PDF file if and only if you use ADOBE ACROBAT
to convert your PS file to PDF.
%
Check that you have set the paper size to A4 (and NOT to letter) in your
dvi2ps converter, in Adobe Acrobat if you use it, and in any printer driver
that you could use.  You also have to disable the 'scale to fit paper' option
of your printer driver.
%
In any case, please check carefully that the final size of the top and
bottom margins is 5.2 cm and of the left and right margins is 4.4 cm.
%t is your responsibility to verify this important requirement.  If these margin requirements and not fulfilled at the end of your file generation process, please use the commands at the beginning of the ESANNV2.tex file to correct them.  Otherwise, please do not modify these commands.

\subsection{Additional packages and functions}

Update the sample file according to your text. You can add
packages or declare new \LaTeX\ functions if and only if there is no conflict between your packages and the esannV2.cls style file.

\subsection{Style information}

\subsubsection{Page numbers}
Please do not add page numbers to this style; page numbers will be added by the publisher.
\subsubsection{Page headings}
Do not add headings to your document.
\subsection{Mathematics}
You may include additional packages for typesetting
algorithms, mathematical formula or to define new operators and environments
if and only if there is no conflict with the esannV2.cls
file.

It is recommended to avoid the numbering of equations when not
necessary. When dealing with equation arrays, it could be
necessary to label several (in)equalities. You can do it using the
`$\backslash$stackrel' operator (see the ESANNV2.tex source file);
example:

\begin{eqnarray}
c&=&|d|+|e|\nonumber\\
&\stackrel{\text{(a)}}{=}&d+e\nonumber\\
&\stackrel{\text{(b)}}{\geq}&\sqrt{f}\enspace,
\end{eqnarray}
\noindent where the equality (a) results from the fact that both
$d$ and $e$ are positive while (b) comes from the definition of
$f$.

\subsection{Tables and figures}

Figure \ref{Fig:MV} shows an example of figure and related
caption.  Do not use too small symbols and lettering in your
figures.  Warning: your paper will be printed in black and white
in the proceedings.  You may insert color figures, but it is your
responsibility to check that they print correctly in black and
white.  The color version will be kept in the ESANN electronic
proceedings available on the web.

\begin{figure}[h!]
\centering
\includegraphics[scale=0.6]{ESANN2005BW.eps}
\caption{ESANN 2005: Announcement and call for
papers.}\label{Fig:MV}
\end{figure}

Table \ref{Tab:AgeWeight} shows an example of table.

\begin{table}[h!]
  \centering
  \begin{tabular}{|c|c|c|}
    \hline
    ID & age & weight \\
    \hline
    1& 15 & 65 \\
    2& 24 & 74\\
    3& 18 & 69 \\
    4& 32 & 78 \\
    \hline
  \end{tabular}
  \caption{Age and weight of people.}\label{Tab:AgeWeight}
\end{table}

\section{Citation}
This ESANNV2.tex file defines how to insert references, both for
BiBTeX and non-BiBTeX users.  Please read the instructions in this
file.

% ****************************************************************************
% BIBLIOGRAPHY AREA
% ****************************************************************************

\begin{footnotesize}

% IF YOU DO NOT USE BIBTEX, USE THE FOLLOWING SAMPLE SCHEME FOR THE REFERENCES
% ----------------------------------------------------------------------------
\begin{thebibliography}{99}

% For books
\bibitem{Haykin_book} S. Haykin, editor. \emph{Unsupervised Adaptive Filtering vol.1 : Blind Source Separation}, John Willey ans Sons, New York, 2000.

% For articles
\bibitem{DelfosseLoubaton_article}N. Delfosse and P. Loubaton, Adaptibe blind separation of sources: A deflation
approach, \emph{Signal Processing}, 45:59-83, Elsevier, 1995.

% For paper in proceedings published as serie books (LNCS,...)
\bibitem{CrucCichAmari_bookproceedings} S. Cruces, A. Cichocki and S. Amari, The minimum entropy and cumulants based contrast functions for blind source extraction. In J. Mira and A. Prieto, editors, proceedings of the 6$^{th}$ \emph{international workshop on artificial neural networks} ({IWANN} 2001), Lecture Notes in Computer Science 2085, pages 786-793,
Springer-Verlag, 2001.

% For paper in conference proceedings
\bibitem{VrinsArchambeau_proceedings} F. Vrins, C. Archambeau and M. Verleysen, Towards a local separation performances estimator using common ICA contrast functions? In M. Verleysen, editor, \emph{proceedings of the $12^{th}$
European Symposium on Artificial Neural Networks} ({ESANN} 2004),
d-side pub., pages 211-216, April 28-30, Bruges (Belgium), 2004.

% For Technical Report
\bibitem{Stone_TechRep} J. V. Stone and J. Porrill, Undercomplete independent component analysis for signal separation and dimension
reduction. Technical Report, Psychology Department, Sheffield
University, Sheffield, S10 2UR, England, October 1997.
\end{thebibliography}
% ----------------------------------------------------------------------------

% IF YOU USE BIBTEX,
% - DELETE THE TEXT BETWEEN THE TWO ABOVE DASHED LINES
% - UNCOMMENT THE NEXT TWO LINES AND REPLACE 'Name_Of_Your_BibFile'

%\bibliographystyle{unsrt}
%\bibliography{Name_Of_Your_BibFile}

\end{footnotesize}

% ****************************************************************************
% END OF BIBLIOGRAPHY AREA
% ****************************************************************************

\end{document}

### 2017-11-16 ESSAN/esannV2.tex
Draft

--- 2017-11-16 ESSAN/article.tex
@@
 \usepackage{amssymb,amsmath,array}
-\graphicspath{{images/}}
 
@@
 %***********************************************************************
-\author{Ram\'on H. Mart\'inez-Mayorquin $^1$, Anders Lansner and $^2$ Pawel Herman $^3$
+\author{First author$^1$ and Second author$^2$
 %
 % Optional short acknowledgment: remove next line if non-needed
-% \thanks{This is an optional funding source acknowledgement.}
+\thanks{This is an optional funding source acknowledgement.}
 %
@@
 % Addresses and institutions (remove "1- " in case of a single institution)
-1- KTH - Dept of First Author \\
+1- School of First Author - Dept of First Author \\
 Address of First Author's school - Country of First Author's
@@
 \vspace{.1cm}\\
-2- KTH - Dept of Second Author \\
+2- School of Second Author - Dept of Second Author \\
 Address of Second Author's school - Country of Second Author's school
@@
 }
-
 %***********************************************************************
@@
 
-\section{Introduction}
+\section{Typesetting an ESANN document using \LaTeX}
 
@@
 
-\subsection{Neural Network}
-
-\begin{align}
-\tau_m \dfrac{dx}{dt} &= \Phi(wx - \theta) \\
-\tau_z \dfrac{dz}{dt} &= x - z
-\end{align}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=2.0]{diagram.eps}
-\caption{ESANN 2005: Announcement and call for
-papers.}\label{Fig:MV}
-\end{figure}
-
+\subsection{Page format and margins}
 Please avoid using DVI2PDF or PS2PDF converters: some undesired

### 2017-11-16 network.py

@@ -25,9 +25,9 @@ def run_network_recall(N, w, G, threshold, tau_m, tau_z,  T, dt, I_cue, T_cue, s
         current_history.append(np.copy(current))
         current = np.dot(w, z) + sigma * np.random.randn(N)
         x += (dt/tau_m) * (phi(G, current - threshold) - x)
-        z += (dt/ tau_z) * (x - z)
         if i < steps_cue:
             x[I_cue] = 1
+        z += (dt / tau_z) * (x - z)
 
     x_history = np.array(x_history)
     z_history = np.array(z_history)
@@ -99,5 +99,41 @@ def train_network(N, dt, training_time, inter_sequence_time, sequences, tau_z, t
 
     if save_w_history:
         dic['w_history'] = np.array(w_history)
+
+    return dic
+
+
+def run_network_recall_limit(N, w, G, threshold, tau_m, tau_z,  T, dt, I_cue, T_cue, sigma=0):
+
+    x = np.zeros(N)
+    current = np.zeros(N)
+    z = np.zeros(N)
+    x[I_cue] = 1.0  # Initial condition
+    z[I_cue] = 0.0
+
+    x_history = []
+    z_history = []
+    current_history = []
+
+    steps = int(T / dt)
+    steps_cue = int(T_cue / dt)
+    for i in range(steps):
+        x_history.append(np.copy(x))
+        z_history.append(np.copy(z))
+        current_history.append(np.copy(current))
+        current = np.dot(w, z) + sigma * np.random.randn(N)
+        x = phi(G, current - threshold)
+        if i < steps_cue:
+            x[I_cue] = 1
+        z += (dt / tau_z) * (x - z)
+
+    x_history = np.array(x_history)
+    z_history = np.array(z_history)
+    current_history = np.array(current_history)
+
+    dic = {}
+    dic['x'] = x_history
+    dic['z'] = z_history
+    dic['current'] = current_history
 
     return dic

### 2017-11-16 plot_producers/esan_recall.py

@@ -7,7 +7,7 @@ import numpy as np
 import seaborn as sns
 from mpl_toolkits.axes_grid1 import make_axes_locatable
 
-plt.rcParams["figure.figsize"] = [16,9]
+plt.rcParams["figure.figsize"] = [16, 9]
 sns.set(font_scale=3.0)
 
 from network import train_network, run_network_recall
@@ -25,7 +25,7 @@ def grayify_cmap(cmap):
     luminance = np.sqrt(np.dot(colors[:, :3] ** 2, RGB_weight))
     colors[:, :3] = luminance[:, np.newaxis]
 
-    return cmap.from_list(cmap.name + "_grayscale", colors, cmap.N)
+    return matplotlib.colors.LinearSegmentedColormap(cmap.name + "_grayscale", colors, cmap.N)
 
 # First the connectivity matrix
 N = 10
@@ -44,7 +44,7 @@ G = 100.0
 tau_m = 0.010
 T = 2.0
 I_cue = 0
-T_cue = 0.050
+T_cue = 0.100
 dt = 0.001
 tau_z = 0.050
 threshold = 0.3
@@ -68,12 +68,22 @@ patterns = [0, 1, 2, 3, 4, 5, 6, 7, 8, 9]
 # patterns = sequence
 
 norm = matplotlib.colors.Normalize(0, N)
-cmap = 'inferno'
+cmap = matplotlib.cm.inferno
 
 for pattern in patterns:
-    ax1.plot(time, x_history[:, pattern], color=norm(pattern), label='x' + str(pattern))
-    ax2.plot(time, current_history[:, pattern], label='current' + str(pattern))
-    ax3.plot(time, z_history[:, pattern], label='z ' + str(pattern))
+    width = pattern * 0.5 + 0.5
+    if pattern == 0:
+        label = 'Cue'
+    else:
+        label = str(pattern)
+
+    ax1.plot(time, x_history[:, pattern], color=cmap(norm(pattern)), linewidth=width, label=label)
+    ax2.plot(time, current_history[:, pattern], color=cmap(norm(pattern)), linewidth=width, label='x' + label)
+    ax3.plot(time, z_history[:, pattern], color=cmap(norm(pattern)), linewidth=width, label='x' + label)
+
+ax1.grid()
+ax2.grid()
+ax3.grid()
 
 ax1.set_ylim([-0.1, 1.1])
[…]

### 2017-11-24 analysis.py
Commit

@@ -68,3 +68,39 @@ def time_t2_local(tau_z, A, T, I, threshold):
     B = np.exp(-t1 / tau_z)
 
     return tau_z * np.log((A * B - I) / (A - (threshold + I)))
+
+
+def create_sequence_chain(number_of_sequences, half_width, units_to_overload):
+    chain = []
+    number = 0
+    for _ in range(number_of_sequences):
+
+        sequence = []
+
+        # The first half
+        i = 0
+        while i < half_width:
+            if number in units_to_overload:
+                number += 1
+
+            else:
+                sequence.append(number)
+                number += 1
+                i += 1
+
+        # The overload units in the middle
+        sequence += units_to_overload
+
+        # The second half
+        i = 0
+        while i < half_width:
+            if number in units_to_overload:
+                number += 1
+            else:
+                sequence.append(number)
+                number += 1
+                i += 1
+
+        chain.append(sequence)
+
+    return chain

### 2017-11-24 connectivity.py

@@ -1,8 +1,8 @@
 import numpy as np
 
 
-def designed_matrix_sequences(N, sequences, self_excitation=1, transition=1, inhbition=1):
-    w = np.ones((N, N)) * -inhbition
+def designed_matrix_sequences(N, sequences, self_excitation=1, transition=1, inhibition=1):
+    w = np.ones((N, N)) * -inhibition
 
     # Self-excitation
     for sequence in sequences:
@@ -17,7 +17,7 @@ def designed_matrix_sequences(N, sequences, self_excitation=1, transition=1, inh
     return w
 
 
-def designed_matrix_sequences_local(N, sequences, self_excitation=1, transition=1, inhbition=1):
+def designed_matrix_sequences_local(N, sequences, self_excitation=1, transition=1, inhibition=1):
     w = np.zeros((N, N))
 
     # Self-excitation
@@ -29,6 +29,6 @@ def designed_matrix_sequences_local(N, sequences, self_excitation=1, transition=
     for sequence in sequences:
         for index in range(len(sequence) - 1):
             w[sequence[index + 1], sequence[index]] = transition
-            w[sequence[index], sequence[index + 1]] = -inhbition
+            w[sequence[index], sequence[index + 1]] = -inhibition
 
     return w

### 2017-11-24 play.py

@@ -1,0 +1,76 @@
+import sklearn
+import numpy as np
+import sklearn.linear_model as linear_model
+
+from network import run_network_recall, train_network, run_network_recall_limit
+from connectivity import designed_matrix_sequences, designed_matrix_sequences_local
+from analysis import get_recall_duration_for_pattern, get_recall_duration_sequence
+from analysis import time_t1, time_t2, time_t1_local, time_t2_local, time_t2_complicated
+
+from sklearn.model_selection import train_test_split
+
+N = 10
+tau_z = 0.150
+tau_z_post = 0.005
+tau_w = 0.100
+max_w = 30.0
+min_w = -3.0
+
+training_time = 0.100
+inter_sequence_time = 1.000
+dt = 0.001
+epochs = 4
+sequence1 = [0, 1, 2, 3, 4, 5, 6, 7, 8, 9]
+sequences = [sequence1]
+
+pattern = 3
+pattern_from = 2
+
+saved_matrix = True
+outfile = './data1.npz'
+if not saved_matrix:
+    n_samples = 5000
+    max_w_vector = np.random.uniform(low=1, high=200, size=n_samples)
+    min_w_vector = -np.random.uniform(low=1, high=100, size=n_samples)
+    tau_z_vector = np.random.uniform(low=0.050, high=1.050, size=n_samples)
+
+    X = np.zeros((n_samples, 3))
+    y = np.zeros((n_samples, 3))
+
+    for index, (tau_z, max_w, min_w) in enumerate(zip(tau_z_vector, max_w_vector, min_w_vector)):
+        dic = train_network(N, dt, training_time, inter_sequence_time, sequences, tau_z, tau_z_post,
+                        tau_w, epochs=epochs, max_w=max_w, min_w=min_w, save_w_history=True, pre_rule=True)
+
+
+        w = dic['w']
+        self = w[pattern, pattern]
+        exc = w[pattern, pattern_from]
+        inh = w[pattern_from, pattern]
+        y[index, 0] = self
+        y[index, 1] = exc
+        y[index, 2] = inh
+
+        X[index, 0] = tau_z
+        X[index, 1] = max_w
+        X[index, 2] = min_w
+
+
+
+    # Now we save both X and Y
+
+    np.savez(outfile, x=X, y=y)
+else:
+    npz_file = np.load(outfile)
+    X = npz_file['x']
+    y = npz_file['y']
+
+# Here we do the classifier
[…]

### 2017-11-25 plot_producers/essan_training_rule.py
The first learning rule plot made

@@ -1,0 +1,102 @@
+import sys
+sys.path.append('../')
+import matplotlib
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+import numpy as np
+import seaborn as sns
+from mpl_toolkits.axes_grid1 import make_axes_locatable
+
+plt.rcParams["figure.figsize"] = [16, 9]
+sns.set(font_scale=3.0)
+
+from network import run_network_recall, train_network, run_network_recall_limit
+
+N = 10
+tau_z = 0.150
+tau_z_1 = 0.050
+tau_z_post = 0.005
+tau_w = 0.100
+max_w = 30.0
+min_w = -3.0
+
+training_time = 0.100
+inter_sequence_time = 1.000
+dt = 0.001
+epochs = 10
+sequence1 = [0, 1, 2, 3, 4, 5, 6, 7, 8, 9]
+sequences = [sequence1]
+
+dic = train_network(N, dt, training_time, inter_sequence_time, sequences, tau_z, tau_z_post,
+                    tau_w, epochs=epochs, max_w=max_w, min_w=min_w, save_w_history=True)
+
+w = dic['w']
+x_total = dic['x']
+z_history = dic['z']
+z_post_history = dic['z_post']
+w_history = dic['w_history']
+
+dic = train_network(N, dt, training_time, inter_sequence_time, sequences, tau_z_1, tau_z_post,
+                    tau_w, epochs=epochs, max_w=max_w, min_w=min_w, save_w_history=True)
+
+w1 = dic['w']
+
+epoch_length = N * (training_time) + (inter_sequence_time )
+T = epoch_length * epochs
+time = np.arange(0, T + dt, dt)
+
+w_10 = w_history[:, 1, 0]
+w_01 = w_history[:, 0, 1]
+w_11 = w_history[:, 1, 1]
+
+fig = plt.figure(figsize=(16, 12))
+gs = gridspec.GridSpec(2, 2)
+ax1 = fig.add_subplot(gs[0, 0])
+ax2 = fig.add_subplot(gs[1, :])
+ax3 = fig.add_subplot(gs[0, 1])
+
+norm = matplotlib.colors.Normalize(0, 2)
+cmap = matplotlib.cm.inferno
+
+# The first w plot
+im = ax1.matshow(w, cmap=cmap)
+divider = make_axes_locatable(ax1)
[…]

### 2017-11-25 network.py
Modifications to the learning frunction

@@ -42,7 +42,7 @@ def run_network_recall(N, w, G, threshold, tau_m, tau_z,  T, dt, I_cue, T_cue, s
 
 
 def train_network(N, dt, training_time, inter_sequence_time, sequences, tau_z, tau_z_post, tau_w,
-                  epochs=1, max_w=1.0, min_w=None, save_w_history=False):
+                  epochs=1, max_w=1.0, min_w=None, save_w_history=False, pre_rule=True):
 
     w = np.zeros((N, N))
     w_history = [w]
@@ -81,7 +81,10 @@ def train_network(N, dt, training_time, inter_sequence_time, sequences, tau_z, t
 
         normal = np.outer(z_post, z)
         # This is the pre-synaptic rule (check this statement)
-        negative = np.outer(1 - z_post, z)
+        if pre_rule:
+            negative = np.outer(1 - z_post, z)
+        else:
+            negative = np.outer(1 - z, z_post)
         if min_w is None:
             w += (dt / tau_w) * ((max_w - w) * normal - negative)
         else:

### 2017-11-25 plot_producers/essan_dynamica_range.py

@@ -1,0 +1,70 @@
+import sys
+sys.path.append('../')
+import matplotlib
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+import numpy as np
+import seaborn as sns
+from mpl_toolkits.axes_grid1 import make_axes_locatable
+
+plt.rcParams["figure.figsize"] = [16, 9]
+sns.set(font_scale=3.0)
+
+from network import run_network_recall, train_network, run_network_recall_limit
+from connectivity import designed_matrix_sequences, designed_matrix_sequences_local
+from analysis import get_recall_duration_for_pattern, get_recall_duration_sequence
+from analysis import time_t1, time_t2, time_t1_local, time_t2_local, time_t2_complicated
+
+N = 10
+sequences = [[0, 1, 2, 3, 4, 5, 6, 7, 8, 9]]
+self_excitation = 3.0
+transition = 0.5
+inhibition = 50.0
+
+threshold = 0.5
+tau_z = 0.050
+
+G = 100.0
+tau_m = 0.010
+T = 20.0
+I_cue = 0
+T_cue = 0.100
+dt = 0.001
+
+pattern = 2
+transition_vector_2 = np.arange(0.55, 1.0, 0.05)
+transition_vector_1 = np.logspace(-14, -0.3, 20) + 0.5
+transition_vector = np.hstack((transition_vector_1, transition_vector_2))
+transition_vector.sort()
+
+recall_times = []
+for transition in transition_vector:
+    w = designed_matrix_sequences(N, sequences, self_excitation=self_excitation, transition=transition,
+                              inhibition=inhibition)
+
+    dic = run_network_recall_limit(N, w, G, threshold, tau_m, tau_z,  T, dt, I_cue, T_cue)
+    x_history = dic['x']
+    duration = get_recall_duration_for_pattern(x_history, pattern, dt)
+    recall_times.append(duration)
+
+
+# The figure
+time_t1_recall = time_t1(tau_z, T=transition_vector, I=inhibition, threshold=threshold)
+
+fig = plt.figure()
+ax = fig.add_subplot(111)
+
+ax.plot(transition_vector, recall_times, '*-', markersize=13, label='recall time')
+ax.plot(transition_vector, time_t1_recall, label='theoretical')
+
+ax.axhline(0, ls='--', color='black')
+ax.axvline(threshold, ls=':', color='grey', label='threshold')
+
[…]

### 2017-11-25 play.py
Added the plot fo the learnings rule

@@ -26,10 +26,10 @@ sequences = [sequence1]
 pattern = 3
 pattern_from = 2
 
-saved_matrix = True
-outfile = './data1.npz'
+saved_matrix = False
+outfile = './data2.npz'
 if not saved_matrix:
-    n_samples = 5000
+    n_samples = 10000
     max_w_vector = np.random.uniform(low=1, high=200, size=n_samples)
     min_w_vector = -np.random.uniform(low=1, high=100, size=n_samples)
     tau_z_vector = np.random.uniform(low=0.050, high=1.050, size=n_samples)

### 2017-11-25 plot_producers/essan_training_rule.py

@@ -12,9 +12,11 @@ sns.set(font_scale=3.0)
 
 from network import run_network_recall, train_network, run_network_recall_limit
 
+independent_labels = False  # Whehter you want an extra weight trained matrix or the labels
+
 N = 10
-tau_z = 0.150
-tau_z_1 = 0.050
+tau_z = 0.050
+tau_z_1 = 0.150
 tau_z_post = 0.005
 tau_w = 0.100
 max_w = 30.0
@@ -53,7 +55,10 @@ fig = plt.figure(figsize=(16, 12))
 gs = gridspec.GridSpec(2, 2)
 ax1 = fig.add_subplot(gs[0, 0])
 ax2 = fig.add_subplot(gs[1, :])
-ax3 = fig.add_subplot(gs[0, 1])
+
+if not independent_labels:
+    print('here 1')
+    ax3 = fig.add_subplot(gs[0, 1])
 
 norm = matplotlib.colors.Normalize(0, 2)
 cmap = matplotlib.cm.inferno
@@ -70,22 +75,11 @@ ax1.xaxis.set_ticklabels([])
 ax1.yaxis.set_ticklabels([])
 ax1.grid()
 
-# The second w plot
-im3 = ax3.matshow(w1, cmap=cmap)
-divider = make_axes_locatable(ax3)
-cax = divider.append_axes('right', size='5%', pad=0.05)
-fig.colorbar(im3, cax=cax)
-
-ax3.set_xlabel('Pre')
-ax3.set_ylabel('Post')
-ax3.xaxis.set_ticklabels([])
-ax3.yaxis.set_ticklabels([])
-ax3.grid()
 
 # The epochs plot
-ax2.plot(time, w_10, color='blue', ls=':', linewidth=2, label='Exc')
-ax2.plot(time, w_01, color='red', ls='-', linewidth=4, label='Inh')
-ax2.plot(time, w_11, color='green', ls='--', linewidth=6, label='Self')
+ax2.plot(time, w_10,  ':', color='blue', markersize=1, linewidth=2, label='Exc')
+ax2.plot(time, w_01, '-', color='red',  markersize=1, linewidth=4, label='Inh')
+ax2.plot(time, w_11, '--', color='green', markersize=1, linewidth=6, label='Self')
 
 tick_spacing = 2
 labels = ['0', '1', '2', '3', '4', '5', '6', '7', '8', '9', '10']
[…]

### 2017-11-25 plot_producers/essan_training_rule_quantities.py

@@ -1,0 +1,294 @@
+import sys
+sys.path.append('../')
+import matplotlib
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+import numpy as np
+import seaborn as sns
+from mpl_toolkits.axes_grid1 import make_axes_locatable
+
+plt.rcParams["figure.figsize"] = [16, 9]
+sns.set(font_scale=3.0)
+
+from network import run_network_recall, train_network, run_network_recall_limit
+
+legends_instead_of_post = True
+
+
+fig = plt.figure(figsize=(16, 12))
+ax1 = fig.add_subplot(321)
+ax2 = fig.add_subplot(322)
+ax3 = fig.add_subplot(323)
+if not legends_instead_of_post:
+    ax4 = fig.add_subplot(324)
+ax5 = fig.add_subplot(325)
+ax6 = fig.add_subplot(326)
+
+plt.tight_layout()
+
+# Some general parameters
+N = 10
+inter_sequence_time = 1.000
+dt = 0.001
+sequence1 = [0, 1, 2, 3, 4, 5, 6, 7, 8, 9]
+sequences = [sequence1]
+epochs = 5
+pattern = 3
+pattern_from = 2
+pre_rule = True
+
+
+# Training time
+training_times_vector = np.arange(0.050, 1.050, 0.050)
+w_self = []
+w_transition = []
+w_inhition = []
+
+tau_z = 0.050
+tau_z_post = 0.005
+tau_w = 0.100
+max_w = 10.0
+min_w = -3.0
+
+training_time = 0.100
+
+for training_time in training_times_vector:
+
+    dic = train_network(N, dt, training_time, inter_sequence_time, sequences, tau_z,
+                        tau_z_post, tau_w, epochs=epochs, max_w=max_w, min_w=min_w, save_w_history=False, pre_rule=True)
+
+    w = dic['w']
+    w_self.append(w[pattern, pattern])
+    w_transition.append(w[pattern, pattern_from])
+    w_inhition.append(w[pattern_from, pattern])
+
+
+ax1.plot(training_times_vector, w_transition, 'o:', color='blue', linewidth=4, markersize=13, label='Exc')
+ax1.plot(training_times_vector, w_inhition, '^-', color='red', linewidth=6, markersize=18, label='Inh')
[…]

### 2017-11-25 ESSAN/article.tex
new working example and dynamical range plots

@@
 
-\subsection{Neural Network}
+\section{The neural network}
 
@@
 \includegraphics[scale=2.0]{diagram.eps}
-\caption{ESANN 2005: Announcement and call for
-papers.}\label{Fig:MV}
+\caption{The connectivity}\label{Fig:diagram}
 \end{figure}
 
-Please avoid using DVI2PDF or PS2PDF converters: some undesired
-shifting/scaling may occur when using these programs
-It is strongly recommended to use the DVIPS converters, and to submit
-PS file. You may submit a PDF file if and only if you use ADOBE ACROBAT
-to convert your PS file to PDF.
-%
-Check that you have set the paper size to A4 (and NOT to letter) in your
-dvi2ps converter, in Adobe Acrobat if you use it, and in any printer driver
-that you could use.  You also have to disable the 'scale to fit paper' option
-of your printer driver.
-%
-In any case, please check carefully that the final size of the top and
-bottom margins is 5.2 cm and of the left and right margins is 4.4 cm.
-%t is your responsibility to verify this important requirement.
-If these margin requirements and not fulfilled at the end of your file generation process, please use the commands at the beginning of the ESANNV2.tex file to correct them.
-Otherwise, please do not modify these commands.
-
-\subsection{Additional packages and functions}
-
-Update the sample file according to your text. You can add
-packages or declare new \LaTeX\ functions if and only if there is no conflict between your packages and the esannV2.cls style file.
-
-\subsection{Style information}
-
-\subsubsection{Page numbers}
-Please do not add page numbers to this style; page numbers will be added by the publisher.
-\subsubsection{Page headings}
-Do not add headings to your document.
-\subsection{Mathematics}
-You may include additional packages for typesetting
-algorithms, mathematical formula or to define new operators and environments
-if and only if there is no conflict with the esannV2.cls
-file.
-
-It is recommended to avoid the numbering of equations when not
-necessary. When dealing with equation arrays, it could be
-necessary to label several (in)equalities. You can do it using the
-`$\backslash$stackrel' operator (see the ESANNV2.tex source file);
-example:
-
-\begin{eqnarray}
-c&=&|d|+|e|\nonumber\\
-&\stackrel{\text{(a)}}{=}&d+e\nonumber\\
-&\stackrel{\text{(b)}}{\geq}&\sqrt{f}\enspace,
-\end{eqnarray}
-\noindent where the equality (a) results from the fact that both
-$d$ and $e$ are positive while (b) comes from the definition of
-$f$.
-
-\subsection{Tables and figures}
-
-Figure \ref{Fig:MV} shows an example of figure and related
-caption.  Do not use too small symbols and lettering in your
-figures.  Warning: your paper will be printed in black and white
-in the proceedings.  You may insert color figures, but it is your
-responsibility to check that they print correctly in black and
-white.  The color version will be kept in the ESANN electronic
-proceedings available on the web.
+\subsection{A recall example}
 
@@
 \centering
-\includegraphics[scale=0.6]{ESANN2005BW.eps}
-\caption{ESANN 2005: Announcement and call for
-papers.}\label{Fig:MV}
+\includegraphics[scale=0.3]{recall.eps}
+\caption{A recall example}\label{Fig:recall}
 \end{figure}
 
-Table \ref{Tab:AgeWeight} shows an example of table.
+\section{Dynamic Range}
+\begin{figure}[h!]
+\centering
+\includegraphics[scale=0.25]{dynamical_range.eps}
+\caption{The dynamica range}\label{Fig:dynamical_range}
+\end{figure}
 
-\begin{table}[h!]
+\section{Learning Rule}
+
+\begin{align}
+\tau_w \dfrac{dw}{dt} &= (w_{max} - w) z_{pre}z_{post} + (w_{min} - w) (z_{post} - 1)z_{pre} 
+\end{align}
+
+\begin{figure}[h!]
 \centering
-  \begin{tabular}{|c|c|c|}
-    \hline
-    ID & age & weight \\
-    \hline
-    1& 15 & 65 \\
-    2& 24 & 74\\
-    3& 18 & 69 \\
-    4& 32 & 78 \\
-    \hline
-  \end{tabular}
-  \caption{Age and weight of people.}\label{Tab:AgeWeight}
-\end{table}
+\includegraphics[scale=1.3]{plasticity_diagram.eps}
+\caption{This is how plasticity works}\label{Fig:plasticity diagram}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[scale=0.3]{training_rule1.eps}
+\caption{The training in action}\label{Fig:epochs}
+\end{figure}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[scale=0.3]{training_rule_quantities.eps}
+\caption{The dynamics of the learning rule}\label{Fig:learning_quantities}
+\end{figure}
+
+\section{Disambiguation}
+
+\section{Discussion}
 

### 2017-11-25 play.py

@@ -26,7 +26,7 @@ sequences = [sequence1]
 pattern = 3
 pattern_from = 2
 
-saved_matrix = False
+saved_matrix = True
 outfile = './data2.npz'
 if not saved_matrix:
     n_samples = 10000
@@ -46,6 +46,7 @@ if not saved_matrix:
         self = w[pattern, pattern]
         exc = w[pattern, pattern_from]
         inh = w[pattern_from, pattern]
+
         y[index, 0] = self
         y[index, 1] = exc
         y[index, 2] = inh
@@ -53,8 +54,6 @@ if not saved_matrix:
         X[index, 0] = tau_z
         X[index, 1] = max_w
         X[index, 2] = min_w
-
-
 
     # Now we save both X and Y
 
@@ -73,4 +72,17 @@ score = reg.score(X_test, y_test)
 A = reg.coef_
 B = np.linalg.inv(A)
 print('score', score)
+
+
+# I want to calculate the values for which the self-excitation and the transition are positive
+positive_A = y[:, 0] > 0
+positive_T = y[:, 1] > 0
+negative_I = y[:, 2] < 0
+
+# Now the values that generated the valures in the correct domain are
+domain_indexes = positive_A * positive_T * negative_I
+X_right = X[domain_indexes, : ]
+
+
+
 

### 2017-11-25 plot_producers/essan_dynamical_range.py

@@ -65,6 +65,6 @@ ax.set_xlabel('Transition')
 ax.set_ylabel('Recall time')
 ax.legend()
 
-fig.savefig('./plot_producers/dynamica_range.eps', frameon=False, dpi=110, bbox_inches='tight')
+fig.savefig('./plot_producers/dynamicaL_range.eps', frameon=False, dpi=110, bbox_inches='tight')
 
 

### 2017-11-25 plot_producers/essan_dynamical_range_double.py

@@ -1,0 +1,117 @@
+import sys
+sys.path.append('../')
+import matplotlib
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+import numpy as np
+import seaborn as sns
+from mpl_toolkits.axes_grid1 import make_axes_locatable
+
+plt.rcParams["figure.figsize"] = [16, 9]
+sns.set(font_scale=3.0)
+
+from network import run_network_recall, train_network, run_network_recall_limit
+from connectivity import designed_matrix_sequences, designed_matrix_sequences_local
+from analysis import get_recall_duration_for_pattern, get_recall_duration_sequence
+from analysis import time_t1, time_t2, time_t1_local, time_t2_local, time_t2_complicated
+
+N = 10
+sequences = [[0, 1, 2, 3, 4, 5, 6, 7, 8, 9]]
+self_excitation = 3.0
+transition = 0.5
+inhibition = 50.0
+
+threshold = 0.5
+tau_z = 0.050
+
+G = 100.0
+tau_m = 0.010
+T = 20.0
+I_cue = 0
+T_cue = 0.100
+dt = 0.001
+
+pattern = 2
+transition_vector_2 = np.arange(0.55, 1.0, 0.05)
+transition_vector_1 = np.logspace(-14, -0.3, 20) + 0.5
+transition_vector = np.hstack((transition_vector_1, transition_vector_2))
+transition_vector.sort()
+
+recall_times = []
+for transition in transition_vector:
+    w = designed_matrix_sequences(N, sequences, self_excitation=self_excitation, transition=transition,
+                              inhibition=inhibition)
+
+    dic = run_network_recall_limit(N, w, G, threshold, tau_m, tau_z,  T, dt, I_cue, T_cue)
+    x_history = dic['x']
+    duration = get_recall_duration_for_pattern(x_history, pattern, dt)
+    recall_times.append(duration)
+
+
+# The figure
+time_t1_recall = time_t1(tau_z, T=transition_vector, I=inhibition, threshold=threshold)
+
+
+
+fig = plt.figure()
+gs = gridspec.GridSpec(2, 2)
+
+ax = fig.add_subplot(gs[:, 0])
+ax1 = fig.add_subplot(gs[0, 1])
+ax2 = fig.add_subplot(gs[1, 1])
+
+plt.tight_layout()
+
[…]

### 2017-11-26 network.py
Modified the plotting functions to remove the background

@@ -140,3 +140,42 @@ def run_network_recall_limit(N, w, G, threshold, tau_m, tau_z,  T, dt, I_cue, T_
     dic['current'] = current_history
 
     return dic
+
+def run_network_recall_limit_end(N, w, G, threshold, tau_m, tau_z,  T, dt, I_cue, I_end, T_cue, sigma=0):
+
+    x = np.zeros(N)
+    current = np.zeros(N)
+    z = np.zeros(N)
+    x[I_cue] = 1.0  # Initial condition
+    z[I_cue] = 0.0
+
+    x_history = []
+    z_history = []
+    current_history = []
+
+    steps = int(T / dt)
+    steps_cue = int(T_cue / dt)
+    for i in range(steps):
+        x_history.append(np.copy(x))
+        z_history.append(np.copy(z))
+        current_history.append(np.copy(current))
+        current = np.dot(w, z) + sigma * np.random.randn(N)
+        x = np.heaviside(current - threshold, 1.0)
+        if i < steps_cue:
+            x[I_cue] = 1
+        z += (dt / tau_z) * (x - z)
+
+        if x[I_end] > 0.5:
+            break
+
+    x_history = np.array(x_history)
+    z_history = np.array(z_history)
+    current_history = np.array(current_history)
+
+    dic = {}
+    dic['x'] = x_history
+    dic['z'] = z_history
+    dic['current'] = current_history
+
+    return dic
+

### 2017-11-26 plot_producers/esan_recall.py

@@ -9,6 +9,7 @@ from mpl_toolkits.axes_grid1 import make_axes_locatable
 
 plt.rcParams["figure.figsize"] = [16, 9]
 sns.set(font_scale=3.0)
+sns.set_style(style='white')
 
 from network import train_network, run_network_recall
 from connectivity import designed_matrix_sequences
@@ -37,7 +38,7 @@ sequence1 = [0, 1, 2, 3, 4, 5, 6, 7, 8, 9]
 sequences = [sequence1]
 
 w = designed_matrix_sequences(N, sequences=sequences, self_excitation=self_excitation,
-                              inhbition=inhibition, transition=transition)
+                              inhibition=inhibition, transition=transition)
 
 # Now the recall
 G = 100.0
@@ -64,11 +65,15 @@ ax1 = fig.add_subplot(gs[0, 0])
 ax2 = fig.add_subplot(gs[1, 0])
 ax3 = fig.add_subplot(gs[2, 0])
 
+ax1.grid()
+ax2.grid()
+ax3.grid()
+
 patterns = [0, 1, 2, 3, 4, 5, 6, 7, 8, 9]
 # patterns = sequence
 
 norm = matplotlib.colors.Normalize(0, N)
-cmap = matplotlib.cm.inferno
+cmap = matplotlib.cm.inferno_r
 
 for pattern in patterns:
     width = pattern * 0.5 + 0.5
@@ -102,8 +107,10 @@ ax3.set_xlabel('Time (s)')
 
 # Here we plot our connectivity matrix
 ax_conn = fig.add_subplot(gs[:2, 1])
+ax_conn.grid()
 cmap = 'seismic'
-cmap = matplotlib.cm.inferno
+cmap = matplotlib.cm.inferno_r
+# cmap = matplotlib.cm.gray_r
 # cmap = grayify_cmap(cmap)
 im = ax_conn.imshow(w, cmap=cmap)
 
@@ -117,18 +124,10 @@ divider = make_axes_locatable(ax_conn)
 cax = divider.append_axes('right', size='5%', pad=0.05)
 
 # Let's define our own color map
-print(cmap(0))
-print(cmap(1))
-print(cmap(2))
-print(cmap(3))
-cmap = matplotlib.colors.ListedColormap([cmap(0), cmap(1), cmap(2)])
-bounds = [-3, 0, 1, 2]
-print(cmap(0))
-print(cmap(1))
-print(cmap(2))
-print(cmap(3))
[…]

### 2017-11-26 plot_producers/essan_dynamical_range_double.py

@@ -9,6 +9,7 @@ from mpl_toolkits.axes_grid1 import make_axes_locatable
 
 plt.rcParams["figure.figsize"] = [16, 9]
 sns.set(font_scale=3.0)
+sns.set_style(style='white')
 
 from network import run_network_recall, train_network, run_network_recall_limit
 from connectivity import designed_matrix_sequences, designed_matrix_sequences_local
@@ -62,15 +63,16 @@ ax2 = fig.add_subplot(gs[1, 1])
 
 plt.tight_layout()
 
-ax.plot(transition_vector, recall_times, '*-', markersize=13, label='recall time')
+ax.plot(transition_vector, recall_times, '*', markersize=15, label='recall time')
 ax.plot(transition_vector, time_t1_recall, label='theoretical')
 
 ax.axhline(0, ls='--', color='black')
 ax.axvline(threshold, ls=':', color='grey', label='threshold')
 
-ax.set_title('Dynamic Range')
-ax.set_xlabel('Recall time')
-ax.set_ylabel('Transition')
+ax.set_title('Dynamical Range')
+ax.set_xlabel(r'$Exc_T$')
+ax.set_ylabel('Recall time (s)')
+
 ax.legend()
 
 # Here the examples
@@ -90,12 +92,12 @@ x_history = dic['x']
 
 im1 = ax1.imshow(x_history.T, aspect='auto', cmap=cmap)
 
-ax1.set_xlabel('Unit')
-ax1.set_ylabel('Time')
-ax1.set_title('Transition = ' + str(transition))
+ax1.set_xlabel('Time')
+ax1.set_ylabel('Unit')
+ax1.set_title(r'$Exc_T$ = ' + str(transition))
 ax1.xaxis.set_ticklabels([])
 ax1.yaxis.set_ticklabels([])
-ax1.grid()
+
 
 
 transition = 0.51
@@ -107,11 +109,11 @@ x_history = dic['x']
 # The second example
 im2 = ax2.imshow(x_history.T, aspect='auto', cmap=cmap)
 
-ax2.set_xlabel('Unit')
-ax2.set_ylabel('Time')
-ax2.set_title('Transition = ' + str(transition))
+ax2.set_xlabel('Time')
+ax2.set_ylabel('Unit')
+ax2.set_title(r'$Exc_T$ = ' + str(transition))
 ax2.xaxis.set_ticklabels([])
 ax2.yaxis.set_ticklabels([])
-ax2.grid()
+
 
[…]

### 2017-11-26 plot_producers/essan_training_rule.py

@@ -9,6 +9,7 @@ from mpl_toolkits.axes_grid1 import make_axes_locatable
 
 plt.rcParams["figure.figsize"] = [16, 9]
 sns.set(font_scale=3.0)
+sns.set_style(style='white')
 
 from network import run_network_recall, train_network, run_network_recall_limit
 
@@ -61,7 +62,7 @@ if not independent_labels:
     ax3 = fig.add_subplot(gs[0, 1])
 
 norm = matplotlib.colors.Normalize(0, 2)
-cmap = matplotlib.cm.inferno
+cmap = matplotlib.cm.inferno_r
 
 # The first w plot
 im = ax1.matshow(w, cmap=cmap)
@@ -73,13 +74,13 @@ ax1.set_xlabel('Pre')
 ax1.set_ylabel('Post')
 ax1.xaxis.set_ticklabels([])
 ax1.yaxis.set_ticklabels([])
-ax1.grid()
+
 
 
 # The epochs plot
-ax2.plot(time, w_10,  ':', color='blue', markersize=1, linewidth=2, label='Exc')
-ax2.plot(time, w_01, '-', color='red',  markersize=1, linewidth=4, label='Inh')
-ax2.plot(time, w_11, '--', color='green', markersize=1, linewidth=6, label='Self')
+ax2.plot(time, w_10,  ':', color='blue', markersize=1, linewidth=2, label=r'$Exc_T$')
+ax2.plot(time, w_01, '-', color='red',  markersize=1, linewidth=4, label=r'$Inh$')
+ax2.plot(time, w_11, '--', color='green', markersize=1, linewidth=6, label=r'$Exc_{self}$')
 
 tick_spacing = 2
 labels = ['0', '1', '2', '3', '4', '5', '6', '7', '8', '9', '10']
@@ -107,7 +108,6 @@ if not independent_labels:
     ax3.set_ylabel('Post')
     ax3.xaxis.set_ticklabels([])
     ax3.yaxis.set_ticklabels([])
-    ax3.grid()
 
     handles, labels = ax2.get_legend_handles_labels()
     fig.legend(handles=handles, labels=labels, loc=(0.3, 0.15), fancybox=False, frameon=False, fontsize=28, ncol=3)
@@ -118,4 +118,4 @@ else:
 
 
 
-fig.savefig('./plot_producers/training_rule1.eps', frameon=False, dpi=110, bbox_inches='tight')
+fig.savefig('./plot_producers/training_rule.eps', frameon=False, dpi=110, bbox_inches='tight')

### 2017-11-26 plot_producers/essan_training_rule_quantities.py

@@ -9,10 +9,13 @@ from mpl_toolkits.axes_grid1 import make_axes_locatable
 
 plt.rcParams["figure.figsize"] = [16, 9]
 sns.set(font_scale=3.0)
+sns.set_style(style='white')
+sns.set_style(style='white')
 
 from network import run_network_recall, train_network, run_network_recall_limit
 
 legends_instead_of_post = True
+add_max_and_min_markers = True
 
 
 fig = plt.figure(figsize=(16, 12))
@@ -63,9 +66,13 @@ for training_time in training_times_vector:
     w_inhition.append(w[pattern_from, pattern])
 
 
-ax1.plot(training_times_vector, w_transition, 'o:', color='blue', linewidth=4, markersize=13, label='Exc')
-ax1.plot(training_times_vector, w_inhition, '^-', color='red', linewidth=6, markersize=18, label='Inh')
-ax1.plot(training_times_vector, w_self, '*--', color='green', linewidth=2, markersize=15, label='Self')
+ax1.plot(training_times_vector, w_transition, 'o:', color='blue', linewidth=4, markersize=13, label=r'$Exc_T$')
+ax1.plot(training_times_vector, w_inhition, '^-', color='red', linewidth=6, markersize=18, label=r'$Inh$')
+ax1.plot(training_times_vector, w_self, '*--', color='green', linewidth=2, markersize=15, label=r'$Exc_{self}$')
+
+if add_max_and_min_markers:
+    ax1.plot(training_times_vector, np.ones(len(w_self)) * max_w, '--', color='gray', label=r'$w_{max}$')
+    ax1.plot(training_times_vector, np.ones(len(w_self)) * min_w, '--', color='gray', label=r'$w_{min}$' )
 
 ax1.axhline(0, ls='--', color='black')
 
@@ -108,6 +115,10 @@ ax2.plot(tau_w_vector, w_transition, 'o:', color='blue', linewidth=4, markersize
 ax2.plot(tau_w_vector, w_inhition, '^-', color='red', linewidth=6, markersize=18, label='Inh')
 ax2.plot(tau_w_vector, w_self, '*--', color='green', linewidth=2, markersize=15, label='Self')
 
+if add_max_and_min_markers:
[…]

### 2017-11-28 network.py
Integral analysis performed

@@ -141,6 +141,7 @@ def run_network_recall_limit(N, w, G, threshold, tau_m, tau_z,  T, dt, I_cue, T_
 
     return dic
 
+
 def run_network_recall_limit_end(N, w, G, threshold, tau_m, tau_z,  T, dt, I_cue, I_end, T_cue, sigma=0):
 
     x = np.zeros(N)

### 2017-11-28 play.py

@@ -2,7 +2,7 @@ import sklearn
 import numpy as np
 import sklearn.linear_model as linear_model
 
-from network import run_network_recall, train_network, run_network_recall_limit
+from network import run_network_recall, train_network, run_network_recall_limit, run_network_recall_limit_end
 from connectivity import designed_matrix_sequences, designed_matrix_sequences_local
 from analysis import get_recall_duration_for_pattern, get_recall_duration_sequence
 from analysis import time_t1, time_t2, time_t1_local, time_t2_local, time_t2_complicated
@@ -10,11 +10,9 @@ from analysis import time_t1, time_t2, time_t1_local, time_t2_local, time_t2_com
 from sklearn.model_selection import train_test_split
 
 N = 10
-tau_z = 0.150
 tau_z_post = 0.005
 tau_w = 0.100
-max_w = 30.0
-min_w = -3.0
+
 
 training_time = 0.100
 inter_sequence_time = 1.000
@@ -23,37 +21,62 @@ epochs = 4
 sequence1 = [0, 1, 2, 3, 4, 5, 6, 7, 8, 9]
 sequences = [sequence1]
 
+threshold  = 0.5
+G = 500.0
+tau_m = 0.010
+I_cue = 0
+I_end = 5
+T = 10.0
+
+
 pattern = 3
 pattern_from = 2
 
-saved_matrix = True
-outfile = './data2.npz'
+saved_matrix = False
+outfile = './data_training.npz'
 if not saved_matrix:
     n_samples = 10000
     max_w_vector = np.random.uniform(low=1, high=200, size=n_samples)
     min_w_vector = -np.random.uniform(low=1, high=100, size=n_samples)
     tau_z_vector = np.random.uniform(low=0.050, high=1.050, size=n_samples)
+    training_vector = np.random.uniform(low=0.100, high=1.000, size=n_samples)
 
-    X = np.zeros((n_samples, 3))
-    y = np.zeros((n_samples, 3))
+    X = np.zeros((n_samples, 4))
+    y = np.zeros((n_samples, 4))
 
-    for index, (tau_z, max_w, min_w) in enumerate(zip(tau_z_vector, max_w_vector, min_w_vector)):
+    for index, (tau_z, max_w, min_w, training_time) in enumerate(zip(tau_z_vector, max_w_vector,
+                                                                     min_w_vector, training_vector)):
+
+        T_cue = tau_z
[…]

### 2017-11-28 play.py
Modified the environment

@@ -27,15 +27,19 @@ tau_m = 0.010
 I_cue = 0
 I_end = 5
 T = 10.0
+pre_rule = False
 
 
 pattern = 3
 pattern_from = 2
 
 saved_matrix = False
-outfile = './data_training.npz'
+n_samples = 5000
+outfile = './data_training_n_samples' + str(n_samples) + 'pre_rule_' + str(pre_rule)
+extension = '.npz'
+outfile += extension
 if not saved_matrix:
-    n_samples = 10000
+
     max_w_vector = np.random.uniform(low=1, high=200, size=n_samples)
     min_w_vector = -np.random.uniform(low=1, high=100, size=n_samples)
     tau_z_vector = np.random.uniform(low=0.050, high=1.050, size=n_samples)
@@ -47,9 +51,11 @@ if not saved_matrix:
     for index, (tau_z, max_w, min_w, training_time) in enumerate(zip(tau_z_vector, max_w_vector,
                                                                      min_w_vector, training_vector)):
 
+        print(index)
+
         T_cue = tau_z
         dic = train_network(N, dt, training_time, inter_sequence_time, sequences, tau_z, tau_z_post,
-                        tau_w, epochs=epochs, max_w=max_w, min_w=min_w, save_w_history=True, pre_rule=True)
+                        tau_w, epochs=epochs, max_w=max_w, min_w=min_w, save_w_history=True, pre_rule=pre_rule)
 
         w = dic['w']
         # Extract ther recall time
@@ -74,12 +80,7 @@ if not saved_matrix:
         X[index, 2] = min_w
         X[index, 3] = training_time
 
-
-
-
-
     # Now we save both X and Y
-
     np.savez(outfile, x=X, y=y)
 else:
     npz_file = np.load(outfile)

### 2017-11-28 plot_producers/essan_training_rule_quantities.py

@@ -12,7 +12,7 @@ sns.set(font_scale=3.0)
 sns.set_style(style='white')
 sns.set_style(style='white')
 
-from network import run_network_recall, train_network, run_network_recall_limit
+from network import train_network, run_network_recall_limit
 
 legends_instead_of_post = True
 add_max_and_min_markers = True
@@ -38,7 +38,7 @@ sequences = [sequence1]
 epochs = 5
 pattern = 3
 pattern_from = 2
-pre_rule = True
+pre_rule = False
 
 
 # Training time
@@ -58,12 +58,13 @@ training_time = 0.100
 for training_time in training_times_vector:
 
     dic = train_network(N, dt, training_time, inter_sequence_time, sequences, tau_z,
-                        tau_z_post, tau_w, epochs=epochs, max_w=max_w, min_w=min_w, save_w_history=False, pre_rule=True)
+                        tau_z_post, tau_w, epochs=epochs, max_w=max_w, min_w=min_w, save_w_history=False, pre_rule=pre_rule)
 
     w = dic['w']
     w_self.append(w[pattern, pattern])
     w_transition.append(w[pattern, pattern_from])
     w_inhition.append(w[pattern_from, pattern])
+
 
 
 ax1.plot(training_times_vector, w_transition, 'o:', color='blue', linewidth=4, markersize=13, label=r'$Exc_T$')
@@ -104,7 +105,7 @@ sequences = [sequence1]
 for tau_w in tau_w_vector:
 
     dic = train_network(N, dt, training_time, inter_sequence_time, sequences, tau_z,
-                        tau_z_post, tau_w, epochs=5, max_w=max_w, min_w=min_w, save_w_history=False)
+                        tau_z_post, tau_w, epochs=5, max_w=max_w, min_w=min_w, save_w_history=False, pre_rule=pre_rule)
 
     w = dic['w']
     w_self.append(w[pattern, pattern])
@@ -149,7 +150,7 @@ sequences = [sequence1]
 for tau_z in tau_z_vector:
 
     dic = train_network(N, dt, training_time, inter_sequence_time, sequences, tau_z,
-                        tau_z_post, tau_w, epochs=5, max_w=max_w, min_w=min_w, save_w_history=False, pre_rule=False)
+                        tau_z_post, tau_w, epochs=5, max_w=max_w, min_w=min_w, save_w_history=False, pre_rule=pre_rule)
 
     w = dic['w']
[…]

### 2017-11-29 ESSAN/article.tex
Modified the plotting functions to remove the background
Added captions to some plots

@@
 
-\section{The neural network}
+\section{Methods}
+The neural network is composed of firing rate units with the following properties.
+First the individual units receive current from all the other units, that is, the system is a recurrently connected neural network with all-to-all connectivity.
+All the neurons send current to each other and the accumulated current is then  thresholded $\theta$ by the logistic function $\Phi(x) = \frac{1}{1 + \exp(-Gx)}$.
+This whole process is condensed into equation \ref{eq1}.
+The current then is integrated with a time constant $\tau_m$, the integration time of the unit, to determine the activation of the unit.
+In order to account for the temporal effects we introduce the idea of z-filters that keep information about the activity during a time period $\tau_z$.
+We can see in equation \ref{eq2} that the filter z follows the activity of the unit at any point in time.
 
 \begin{align}
-\tau_m \dfrac{dx}{dt} &= \Phi(wx - \theta) \\
-\tau_z \dfrac{dz}{dt} &= x - z
+\tau_m \dfrac{dx_i}{dt} &= \Phi\Big(\sum_{j} W_{ij} z_j - \theta \Big) - x_i \label{eq1} \\ 
+\tau_z \dfrac{dz_i}{dt} &= x_i - z_i \label{eq2}
 \end{align}
+
+This system allows for the units to keep influencing the other units through $W$ even when they are not active anymore.
+We will see how this is fundamental for the sequence processing capabilities of our system.
+
+The connectivity of the system is as depicted in figure \ref{Fig:diagram}.
+Note that that from each unit emanate two excitatory connections, one that self-reinforces the unit activity $Exc_{self}$ and another that excites the following neuron (contributing to the sequential structure of the activity) $Exc_T$.
+At the same time every unit inhibits all the units except the next one with $Inh$.
+Of special relevance are the backwards inhibitory projections that suppress the unit once the next unit in the sequence is activated
 
@@
 \centering
-\includegraphics[scale=2.0]{diagram.eps}
-\caption{The connectivity}\label{Fig:diagram}
+\includegraphics[scale=1.5]{diagram.eps}
+\caption{In this diagram we can appreciate the overall connectivity of the system.
+Note that the inhibitory connections are only fully depicted for the first unit for the sake of clarity.}\label{Fig:diagram}
 \end{figure}
 
-\subsection{A recall example}
+\section{Results}
+
+\subsection{The neural network can sustain sequential activity}
+
+In order to recall a particular sequence we cue the first element.
+This means that we artificially activate the first unit (clamping) of the sequence for a short time.
+After this initial cue, if the weights are set right, the neural network continues evolving on its own and a particular sequence is recalled.
+In figure \ref{Fig:recall} we can appreciate a successful recall of a sequence.
 
@@
 \centering
-\includegraphics[scale=0.3]{recall.eps}
-\caption{A recall example}\label{Fig:recall}
+\includegraphics[scale=0.20]{recall.eps}
+\caption{A succesful recall.
+a) The activity of the units as it evolves in time.
+b) The currents that each unit receives as the sequence progresses.
+c) the z-filters activations for each unit.
+d) The connectivity matrix $W$.
+The weights of the matrix can be divided on the excitatory connections: the ones that induce transitions $Exc_{T}$ and the ones that stabilize the unit $Exc_{self}$ and the inhibitory connections $Inh$.}\label{Fig:recall}
 \end{figure}
 
-\section{Dynamic Range}
+%It is important to note that the weights of the connectivity matrix determine the particular behavior that the network will develop.
+In the diagonal of the connectivity matrix we have the weights of the excitation connections that keep the unit active once it already became activited $Exc_{self}$.
+That is, they sustain a positive feed-back loop that keeps the unit active.
+Below the diagonal, on the other hand, we have the weights of the connections that make the sequence transition from one state to the next $Exc_{T}$.
+Finally, the rest of the matrix consists of the weights for the inhibitory connections.
+Of special significance are the backward inhibitory projections above the diagonal, their role is to suppress the last pattern of the sequence once the next unit is activated.
+
+It is now instructive to detail at the level of the dynamics how the sequential activity is propagated.
+After the first unit $x_1$ has been activated for long enough due to clamping the filter $z_1$ will start catching up with the activity by increasing its value.
+When this happens the unit corresponding to $x_2$ will receive a growing current with value $Exc_{T} \times z_1$.
+As this process continues there will a point in time (exactly at $Exc_{T}(1 - e^{-\frac{t}{\tau_z}}) - \theta = 0$) when $x_2$ will receive enough current to become activated.
+This, will again will prompt the filter $z_2$ to start catching up initializing the process described again.
+On the other hand as soon as the second filter starts growing $z_2$ the first unit will start receiving an inhibitory current $Inh \times z_2$ due to the backward inhibitory projection, this will in turn suppress the unit $x_1$ allowing the sequence to propagate.
+Finally, once the unit $x_1$ is finally suppressed the excitatory current mediated by $Exc_T$ will be diminished which in most cases would also suppress the activity of $x_2$ if were not for the self-sustaining current $Exc_{self} \times z_2$.
+In short, the term $Exc_T$ ensures the transition of the activity from one unit of the sequence to the next whereas the competition between $Exc_{self}$ and $Inh$ first lock the pattern unit and then finally suppress it allowing the sequence to evolve towards its end without hindrance.
+
+With the knowledge above in mind we can envision the following  failure scenarios for the sequential dynamics.
+First, if the magnitude of $Exc_{self}$ the activity will not be big enough to overcome the threshold and the sequence will not propagate.
+Second, if $Exc_{Self}$ is enough to keep the unit active but $Exc_{T}$ is not big enough the next unit in the sequence will not be activated and the sequence will not propagate.
+If both $Exc_{self}$ is big enough to sustain the activity and $Exc_{T}$ is big enough to propagate it but the inhibitory connection $Inh$ is not enough to suppress the units the result will be that all the units in the sequence will be activated at the end.
+
+\subsection{The neural network possess a wide dynamic range}
+
+A common problem in sequence learning is to scale the time duration of the sequence according to the needs of the required context.
+In terms of a neural network it means that once the dynamics of sequence recalling are stored in the shape of the weights we can control the \textit{recall time} of the sequence with the internal parameters of the network where the recall time is the amount of time that a particular unit is active during a sequence recall.
+In our network there are a two parameters that a priori are strong candidates to regulate the recall time: the transition weight $Exc_T$ by controlling how fast you activate the next unit in the sequence and the threshold $\theta$ by controlling how long it takes for the current to activate the next unit in a sequence.
+Here we show the exact results for $Exc_T$ but the reasoning for $\theta$ is analogous.
+
+In order to calculate the recall time we make the following consideration.
+Once the unit $x_1$ becomes activated the filter $z_1$ starts increasing its value and the unit $x_2$ will be receiving an excitatory current $Exc_T \times z_1$.
+The point in time t at which the this current overcomes the threshold will be the recall time.
+Using that reasoning we can solve the following equation to calculate the recall time: $Exc_{T}(1 - e^{-\frac{t}{\tau_z}}) - \theta = 0$.
+If we do we end up with the following expression: $T_{recall} = \tau_z \log(\frac{Exc_T}{Exc_T - \theta})$.
+This expression tells us that we actually have a singularity at $Exc_T = \theta$, which for our purposes means that can make the recall time arbitrarily by setting this two parameters close enough.
+We can see an illustration of that in \ref{Fig:dynamical_range} where we also show two examples of sequence recall with different $Exc_T$ values and therefore with different recall times.
+
 \begin{figure}[h!]
@@
 \includegraphics[scale=0.25]{dynamical_range.eps}
-\caption{The dynamica range}\label{Fig:dynamical_range}
+\caption{The dynami}\label{Fig:dynamical_range}
 \end{figure}
 
-\section{Learning Rule}
+\subsection{Learning Rule}
 
@@
 \centering
-\includegraphics[scale=0.3]{training_rule1.eps}
+\includegraphics[scale=0.3]{training_rule.eps}
 \caption{The training in action}\label{Fig:epochs}
@@
 
-\section{Disambiguation}
+\textbf{Training time}
+\begin{itemize}
+\item  $Exc_{self}$: The longer the training time the longer the filter is activated with itself and therefore it grows.
+But given that the pre and post filters do not coincide completely in the temporal domain there are negative contributions which prevent this from reaching the saturation value $w_{max}$
+\item  $Inh$: In this case the the negative contributions come from the time that the units are not activated together with their counterparts which is most of the time (except in the brief transition period) and therefore the value saturates quickly to $w_{min}$.
+\item $Exc_{T}$: In this case what matters is the transition time between two 
+\end{itemize}
+
+\subsection{Disambiguation}
 
@@
 
-% IF YOU DO NOT USE BIBTEX, USE THE FOLLOWING SAMPLE SCHEME FOR THE REFERENCES
-% ----------------------------------------------------------------------------
-\begin{thebibliography}{99}
-
-% For books
-\bibitem{Haykin_book} S. Haykin, editor. \emph{Unsupervised Adaptive Filtering vol.1 : Blind Source Separation}, John Willey ans Sons, New York, 2000.
-
-% For articles
-\bibitem{DelfosseLoubaton_article}N. Delfosse and P. Loubaton, Adaptibe blind separation of sources: A deflation
-approach, \emph{Signal Processing}, 45:59-83, Elsevier, 1995.
-
-% For paper in proceedings published as serie books (LNCS,...)
-\bibitem{CrucCichAmari_bookproceedings} S.
-Cruces, A.
-Cichocki and S.
-Amari, The minimum entropy and cumulants based contrast functions for blind source extraction.
-In J.
-Mira and A.
-Prieto, editors, proceedings of the 6$^{th}$ \emph{international workshop on artificial neural networks} ({IWANN} 2001), Lecture Notes in Computer Science 2085, pages 786-793,
-Springer-Verlag, 2001.
-
-% For paper in conference proceedings
-\bibitem{VrinsArchambeau_proceedings} F.
-Vrins, C.
-Archambeau and M.
-Verleysen, Towards a local separation performances estimator using common ICA contrast functions?
-In M.
-Verleysen, editor, \emph{proceedings of the $12^{th}$
-European Symposium on Artificial Neural Networks} ({ESANN} 2004),
-d-side pub., pages 211-216, April 28-30, Bruges (Belgium), 2004.
-
-% For Technical Report
-\bibitem{Stone_TechRep} J. V. Stone and J. Porrill, Undercomplete independent component analysis for signal separation and dimension
-reduction. Technical Report, Psychology Department, Sheffield
-University, Sheffield, S10 2UR, England, October 1997.
-\end{thebibliography}
-% ----------------------------------------------------------------------------
-
 % IF YOU USE BIBTEX,
@@
 
-%\bibliographystyle{unsrt}
-%\bibliography{Name_Of_Your_BibFile}
+\bibliographystyle{unsrt}
+\bibliography{references.bib}
 

### 2017-11-29 plot_producers/essan_dynamical_range_double.py

@@ -52,7 +52,7 @@ for transition in transition_vector:
 # The figure
 time_t1_recall = time_t1(tau_z, T=transition_vector, I=inhibition, threshold=threshold)
 
-
+captions = True
 
 fig = plt.figure()
 gs = gridspec.GridSpec(2, 2)
@@ -61,13 +61,19 @@ ax = fig.add_subplot(gs[:, 0])
 ax1 = fig.add_subplot(gs[0, 1])
 ax2 = fig.add_subplot(gs[1, 1])
 
+if captions:
+    size = 30
+    fig.text(0.05, 0.95, 'a)', size=size)
+    fig.text(0.55, 0.95, 'b)', size=size)
+    fig.text(0.55, 0.50, 'c)', size=size)
+
 plt.tight_layout()
 
 ax.plot(transition_vector, recall_times, '*', markersize=15, label='recall time')
 ax.plot(transition_vector, time_t1_recall, label='theoretical')
 
 ax.axhline(0, ls='--', color='black')
-ax.axvline(threshold, ls=':', color='grey', label='threshold')
+ax.axvline(threshold, ls=':', color='grey', label=r'threshold $\theta$')
 
 ax.set_title('Dynamical Range')
 ax.set_xlabel(r'$Exc_T$')
@@ -98,8 +104,6 @@ ax1.set_title(r'$Exc_T$ = ' + str(transition))
 ax1.xaxis.set_ticklabels([])
 ax1.yaxis.set_ticklabels([])
 
-
-
 transition = 0.51
 w = designed_matrix_sequences(N, sequences, self_excitation=self_excitation, transition=transition,
                               inhibition=inhibition)
@@ -116,4 +120,4 @@ ax2.xaxis.set_ticklabels([])
 ax2.yaxis.set_ticklabels([])
 
 
-fig.savefig('./plot_producers/dynamical_range_double.eps', frameon=False, dpi=110, bbox_inches='tight')
+fig.savefig('./plot_producers/dynamical_range.eps', frameon=False, dpi=110, bbox_inches='tight')

### 2017-11-29 plot_producers/essan_recall.py

@@ -58,16 +58,24 @@ x_history = dic['x']
 z_history = dic['z']
 current_history = dic['current']
 
+captions = True
+annotations = True
+
 gs = gridspec.GridSpec(3, 2)
 fig = plt.figure(figsize=(16, 12))
+
+# Captions
+if captions:
+    size = 25
+    fig.text(0.05, 0.9, 'a)', size=size)
+    fig.text(0.05, 0.60, 'b)', size=size)
+    fig.text(0.05, 0.35, 'c)', size=size)
+    fig.text(0.5, 0.90, 'd)', size=size)
+
 
 ax1 = fig.add_subplot(gs[0, 0])
 ax2 = fig.add_subplot(gs[1, 0])
 ax3 = fig.add_subplot(gs[2, 0])
-
-ax1.grid()
-ax2.grid()
-ax3.grid()
 
 patterns = [0, 1, 2, 3, 4, 5, 6, 7, 8, 9]
 # patterns = sequence
@@ -76,7 +84,7 @@ norm = matplotlib.colors.Normalize(0, N)
 cmap = matplotlib.cm.inferno_r
 
 for pattern in patterns:
-    width = pattern * 0.5 + 0.5
+    width = 10.0 - pattern * 1.0
     if pattern == 0:
         label = 'Cue'
     else:
@@ -86,9 +94,6 @@ for pattern in patterns:
     ax2.plot(time, current_history[:, pattern], color=cmap(norm(pattern)), linewidth=width, label='x' + label)
     ax3.plot(time, z_history[:, pattern], color=cmap(norm(pattern)), linewidth=width, label='x' + label)
 
-ax1.grid()
-ax2.grid()
-ax3.grid()
 
 ax1.set_ylim([-0.1, 1.1])
 ax3.set_ylim([-0.1, 1.1])
@@ -122,6 +127,17 @@ ax_conn.grid()
 
 divider = make_axes_locatable(ax_conn)
 cax = divider.append_axes('right', size='5%', pad=0.05)
+
+if annotations:
+    ax_conn.annotate(r'$Exc_{T}$', xy=(0, 1), xytext=(0, 5),
+                     arrowprops=dict(facecolor='red', shrink=0.15))
+
+    ax_conn.annotate(r'$Inh$', xy=(1.0, 0), xytext=(4, -1),
+                arrowprops=dict(facecolor='red', shrink=0.05))
+
[…]

### 2017-11-29 ESSAN/article.tex
Updated the article

@@
 This expression tells us that we actually have a singularity at $Exc_T = \theta$, which for our purposes means that can make the recall time arbitrarily by setting this two parameters close enough.
-We can see an illustration of that in \ref{Fig:dynamical_range} where we also show two examples of sequence recall with different $Exc_T$ values and therefore with different recall times.
+In our neural network however, there are more currents into play because of the all-to-all connectivity, we can include those corrections and we end up with an expression of the form $T_{recall} = \tau_z \log(\frac{Exc_T + Inh}{Exc_T - \theta})$ which also possess the after-mentioned property.
+We can see an illustration of the wide dynamical range of our system in figure \ref{Fig:dynamical_range} where we also show two examples of sequence recall with different $Exc_T$ values and therefore with different recall times.
 
@@
 \includegraphics[scale=0.25]{dynamical_range.eps}
-\caption{The dynami}\label{Fig:dynamical_range}
+\caption{The network posses a wide dynamical range.
+a) Here we compare the calculated thoretical expression with simulations with our network for different values of $Exc_T$.
+Note the singularity at $\theta$.
+b) The network activity with a short recall time.
+c) The network activity with a longer recall time ($Exc_T$ close to $\theta$).}\label{Fig:dynamical_range}
 \end{figure}

### 2017-11-29 network.py
New notebook for disambiguation

@@ -93,7 +93,6 @@ def train_network(N, dt, training_time, inter_sequence_time, sequences, tau_z, t
         if save_w_history:
             w_history.append(np.copy(w))
 
-
     dic = {}
     dic['w'] = w
     dic['x'] = x_example

### 2017-11-29 plot_producers/essan_training_rule.py

@@ -14,6 +14,7 @@ sns.set_style(style='white')
 from network import run_network_recall, train_network, run_network_recall_limit
 
 independent_labels = False  # Whehter you want an extra weight trained matrix or the labels
+captions = True
 
 N = 10
 tau_z = 0.050
@@ -31,7 +32,7 @@ sequence1 = [0, 1, 2, 3, 4, 5, 6, 7, 8, 9]
 sequences = [sequence1]
 
 dic = train_network(N, dt, training_time, inter_sequence_time, sequences, tau_z, tau_z_post,
-                    tau_w, epochs=epochs, max_w=max_w, min_w=min_w, save_w_history=True)
+                    tau_w, epochs=epochs, max_w=max_w, min_w=min_w, save_w_history=True, pre_rule=True)
 
 w = dic['w']
 x_total = dic['x']
@@ -40,7 +41,7 @@ z_post_history = dic['z_post']
 w_history = dic['w_history']
 
 dic = train_network(N, dt, training_time, inter_sequence_time, sequences, tau_z_1, tau_z_post,
-                    tau_w, epochs=epochs, max_w=max_w, min_w=min_w, save_w_history=True)
+                    tau_w, epochs=epochs, max_w=max_w, min_w=min_w, save_w_history=True, pre_rule=True)
 
 w1 = dic['w']
 
@@ -53,6 +54,14 @@ w_01 = w_history[:, 0, 1]
 w_11 = w_history[:, 1, 1]
 
 fig = plt.figure(figsize=(16, 12))
+
+if captions:
+    size = 25
+    fig.text(0.10, 0.85, 'a)', size=size)
+    fig.text(0.55, 0.85, 'b)', size=size)
+    fig.text(0.10, 0.48, 'c)', size=size)
+
+
 gs = gridspec.GridSpec(2, 2)
 ax1 = fig.add_subplot(gs[0, 0])
 ax2 = fig.add_subplot(gs[1, :])

### 2017-11-29 ESSAN/article.tex
New plot utilities and last version of the article

@@
 The neural network is composed of firing rate units with the following properties.
-First the individual units receive current from all the other units, that is, the system is a recurrently connected neural network with all-to-all connectivity.
-All the neurons send current to each other and the accumulated current is then  thresholded $\theta$ by the logistic function $\Phi(x) = \frac{1}{1 + \exp(-Gx)}$.
-This whole process is condensed into equation \ref{eq1}.
-The current then is integrated with a time constant $\tau_m$, the integration time of the unit, to determine the activation of the unit.
+First the individual units receive current from all the other units, that is, the system is a recurrent neural network with all-to-all connectivity.
+The input current that a unit receives is  is then  thresholded $\theta$ by the logistic function $\Phi(x) = \frac{1}{1 + \exp(-Gx)}$.
+This whole process is condensed into the equation \ref{eq1}.
+The input current is then integrated with a time constant $\tau_m$ to determine the activation of the unit.
 In order to account for the temporal effects we introduce the idea of z-filters that keep information about the activity during a time period $\tau_z$.
@@
 
-This system allows for the units to keep influencing the other units through $W$ even when they are not active anymore.
-We will see how this is fundamental for the sequence processing capabilities of our system.
+This system allows for the units to keep influencing the other units through $W$ even when they are not longer active.
+We will see how this property is fundamental for the sequence processing capabilities of our system.
 
@@
 After this initial cue, if the weights are set right, the neural network continues evolving on its own and a particular sequence is recalled.
-In figure \ref{Fig:recall} we can appreciate a successful recall of a sequence.
+In figure \ref{Fig:recall} a successful recall process is shown.
 
@@
 \subsection{Learning Rule}
+Here we present a learning rule that allows the system to self-organize in the presence of sequential activity, that is, we can equip our network with the capability to learn how to reproduce sequences.
+We present a learning rule in the equations bellow.
+The first term is completly hebbian \cite{hebb2005organization} and its role is to strength the connections between units that are activate together.
+However, in this case, we use the z-filters instead of the activities directly which allow us to bend units in time.
+We include here another filter like the one in equation \ref{eq2} but with a shorter time constant $\tau_{z_{post}}$.
+The second term is called synaptic gating \cite{andrew2003spiking} and it allow us to create inhibitory connections when two units are not activated together.
+We introduce two different variations of this term to explore the functional implications of both.
+Finally both terms are saturated by maximal and minimal values $w_{max}$ and $w_{min}$.
 
 \begin{align}
-\tau_w \dfrac{dw}{dt} &= (w_{max} - w) z_{pre}z_{post} + (w_{min} - w) (z_{post} - 1)z_{pre} 
+\tag{pre-synaptic}
+\tau_w \dfrac{dw}{dt} &= (w_{max} - w) z_{pre}z_{post} + (w_{min} - w) (1 - z_{post})z_{pre} \label{eq:pre-synaptic rule}  \\
+\tag{post-synaptic}
+\tau_w \dfrac{dw}{dt} &= \underbrace{(w_{max} - w)}_{\text{saturation}} \underbrace{z_{pre}z_{post}}_{\text{hebbian}} + \underbrace{(w_{min} - w)}_{\text{saturation}}\underbrace{ (1 - z_{pre})z_{post}}_{\text{synaptic-gating}} \label{eq:post-synaptic rule}
 \end{align}
+
+The nature and the role of both the hebbian and the synaptic gating terms are better explained in the diagram shown in figure \ref{Fig:plasticity diagram}.
+There we can appreciate the the main difference between the two rules is on whether they produce a negative contribution when either $z_{pre}$ and $z_{post}$ is activated.
+In figure \ref{Fig:epochs} we illustrate the results of the training rule
 
@@
 \includegraphics[scale=1.3]{plasticity_diagram.eps}
-\caption{This is how plasticity works}\label{Fig:plasticity diagram}
+\caption{A scheme of how plasticity works.
+a) Hebbian term contribution.
+b) The contribution is negative when the post-synaptic unit is active.
+c) The contribution is negative when the pre-synaptic unit is active.
+d) If units are not active there is no decay of the weights.}\label{Fig:plasticity diagram}
 \end{figure}
@@
 \includegraphics[scale=0.3]{training_rule.eps}
-\caption{The training in action}\label{Fig:epochs}
+\caption{Training with the learning rule.
+a) An example of training with a small value of $tau_z$.
+b) Same procedure as in a but with a longer time constant $\tau_z$.
+c) evolution of weights with the number of epochs.
+The three situations depicted utilize the pre-synaptic learning rule.}\label{Fig:epochs}
 \end{figure}
+
+\subsubsection{Learning rule and control}
 
@@
 \centering
-\includegraphics[scale=0.3]{training_rule_quantities.eps}
-\caption{The dynamics of the learning rule}\label{Fig:learning_quantities}
+\includegraphics[scale=0.3]{training_rule_quantities_pre_rule_True.eps}
+\caption{The dynamics of the learning rule}\label{Fig:learning_quantities_pre}
 \end{figure}
@@
 
+\begin{figure}[h!]
+\centering
+\includegraphics[scale=0.3]{training_rule_quantities_pre_rule_False.eps}
+\caption{The dynamics of the learning rule}\label{Fig:learning_quantities_post}
+\end{figure}
+
 \subsection{Disambiguation}
@@
 \section{Discussion}
-
-\section{Citation}
-This ESANNV2.tex file defines how to insert references, both for
-BiBTeX and non-BiBTeX users.  Please read the instructions in this
-file.
 

### 2017-11-29 plot_producers/essan_training_rule.py

@@ -79,12 +79,30 @@ divider = make_axes_locatable(ax1)
 cax = divider.append_axes('right', size='5%', pad=0.05)
 fig.colorbar(im, cax=cax)
 
+ax1.set_title(r'$\tau_z$ = ' + str(tau_z))
 ax1.set_xlabel('Pre')
 ax1.set_ylabel('Post')
 ax1.xaxis.set_ticklabels([])
 ax1.yaxis.set_ticklabels([])
 
+# Either a matrix or a legend
+if not independent_labels:
 
+    im3 = ax3.matshow(w1, cmap=cmap)
+    divider = make_axes_locatable(ax3)
+    cax = divider.append_axes('right', size='5%', pad=0.05)
+    fig.colorbar(im3, cax=cax)
+
+    ax3.set_title(r'$\tau_z$ = ' + str(tau_z_1))
+    ax3.set_xlabel('Pre')
+    ax3.set_ylabel('Post')
+    ax3.xaxis.set_ticklabels([])
+    ax3.yaxis.set_ticklabels([])
+
+
+if independent_labels:
+    handles, labels = ax2.get_legend_handles_labels()
+    fig.legend(handles=handles, labels=labels, loc=(0.6, 0.65), fancybox=True, frameon=True, fontsize=28, ncol=1)
 
 # The epochs plot
 ax2.plot(time, w_10,  ':', color='blue', markersize=1, linewidth=2, label=r'$Exc_T$')
@@ -97,34 +115,13 @@ ticks =  [0, 2,    4,   6,   8,   10,  12,  14, 16,  18, 20]
 
 plt.xticks(ticks, labels)
 
-if not independent_labels:
-    print('here 2')
-    # ax2.legend(loc=4)
 
 ax2.set_xlabel('Epochs')
 ax2.set_ylabel('Weight')
 # ax2.set_xlim([-1, 25])
 
-# Either a matrix or a legend
 if not independent_labels:
-    print('here 3')
-    im3 = ax3.matshow(w1, cmap=cmap)
-    divider = make_axes_locatable(ax3)
-    cax = divider.append_axes('right', size='5%', pad=0.05)
-    fig.colorbar(im3, cax=cax)
-
-    ax3.set_xlabel('Pre')
-    ax3.set_ylabel('Post')
-    ax3.xaxis.set_ticklabels([])
-    ax3.yaxis.set_ticklabels([])
-
     handles, labels = ax2.get_legend_handles_labels()
[…]

### 2017-11-29 plot_producers/essan_training_rule_quantities.py

@@ -16,6 +16,8 @@ from network import train_network, run_network_recall_limit
 
 legends_instead_of_post = True
 add_max_and_min_markers = True
+captions = True
+pre_rule = True
 
 
 fig = plt.figure(figsize=(16, 12))
@@ -26,6 +28,15 @@ if not legends_instead_of_post:
     ax4 = fig.add_subplot(324)
 ax5 = fig.add_subplot(325)
 ax6 = fig.add_subplot(326)
+
+if captions:
+    size = 25
+    fig.text(0.05, 0.97, 'a)', size=size)
+    fig.text(0.50, 0.97, 'b)', size=size)
+    fig.text(0.05, 0.65, 'c)', size=size)
+    fig.text(0.05, 0.32, 'd)', size=size)
+    fig.text(0.50, 0.32, 'f)', size=size)
+
 
 plt.tight_layout()
 
@@ -38,7 +49,7 @@ sequences = [sequence1]
 epochs = 5
 pattern = 3
 pattern_from = 2
-pre_rule = False
+
 
 
 # Training time
@@ -96,7 +107,7 @@ tau_w = 0.100
 max_w = 10.0
 min_w = -3.0
 
-training_time = 0.100
+training_time = 0.300
 inter_sequence_time = 1.000
 dt = 0.001
 sequence1 = [0, 1, 2, 3, 4, 5, 6, 7, 8, 9]
@@ -141,7 +152,7 @@ tau_w = 0.100
 max_w = 80.0
 min_w = -50.0
 
-training_time = 0.350
+training_time = 0.500
 inter_sequence_time = 1.000
 dt = 0.001
 sequence1 = [0, 1, 2, 3, 4, 5, 6, 7, 8, 9]
@@ -182,7 +193,7 @@ if not legends_instead_of_post:
     pattern_from = 2
 
     N = 10
-    tau_z = 0.300
+    tau_z = 0.500
     tau_z_post = 0.005
     tau_w = 0.100
     max_w = 100.0
@@ -230,7 +241,7 @@ tau_w = 0.100
 max_w = 30.0
 min_w = -3.0
 
-training_time = 0.100
+training_time = 0.500
 inter_sequence_time = 1.000
 dt = 0.001
 sequence1 = [0, 1, 2, 3, 4, 5, 6, 7, 8, 9]
@@ -276,7 +287,7 @@ tau_w = 0.100
 max_w = 30.0
 min_w = -3.0
 
-training_time = 0.100
+training_time = 0.500
 inter_sequence_time = 1.000
[…]

### 2017-11-29 plot_producers/essan_recall.py
Modified plot recall and diagram

@@ -62,15 +62,17 @@ captions = True
 annotations = True
 
 gs = gridspec.GridSpec(3, 2)
-fig = plt.figure(figsize=(16, 12))
+fig = plt.figure(figsize=(28, 12))
 
 # Captions
 if captions:
-    size = 25
-    fig.text(0.05, 0.9, 'a)', size=size)
-    fig.text(0.05, 0.60, 'b)', size=size)
-    fig.text(0.05, 0.35, 'c)', size=size)
+    size = 35
+    aux_x = 0.08
+    fig.text(aux_x, 0.9, 'a)', size=size)
+    fig.text(aux_x, 0.60, 'b)', size=size)
+    fig.text(aux_x, 0.35, 'c)', size=size)
     fig.text(0.5, 0.90, 'd)', size=size)
+    fig.text(0.5, 0.40, 'e)', size=size)
 
 
 ax1 = fig.add_subplot(gs[0, 0])
@@ -102,16 +104,28 @@ ax1.axhline(0, ls='--', color='black')
 ax2.axhline(0, ls='--', color='black')
 ax3.axhline(0, ls='--', color='black')
 
-ax1.set_ylabel('Unit Activity')
-ax2.set_ylabel('Current')
-ax3.set_ylabel('Z-filters')
+ax1.set_xlim([-0.1, 1.8])
+ax2.set_xlim([-0.1, 1.8])
+ax3.set_xlim([-0.1, 1.8])
+
+
+if False:
+    ax1.set_xlabel('unit activities, x')
+    ax2.set_ylabel('currents')
+    ax3.set_ylabel('z-filters')
+else:
+    ax1.set_title('unit activities, x')
+    ax2.set_title(r'currents, $\Phi$ argument')
+    ax3.set_title('z-filters')
 
 ax1.get_xaxis().set_visible(False)
 ax2.get_xaxis().set_visible(False)
 ax3.set_xlabel('Time (s)')
 
 # Here we plot our connectivity matrix
-ax_conn = fig.add_subplot(gs[:2, 1])
+rect = [0.42, 0.48, 0.40, 0.40]
+# ax_conn = fig.add_subplot(gs[:2, 1])
+ax_conn = fig.add_axes(rect)
 ax_conn.grid()
 cmap = 'seismic'
 cmap = matplotlib.cm.inferno_r
@@ -119,8 +133,8 @@ cmap = matplotlib.cm.inferno_r
 # cmap = grayify_cmap(cmap)
[…]

### 2017-11-29 ESSAN/article.tex
Modified diagrams and plot futher

@@
 \includegraphics[scale=1.5]{diagram.eps}
-\caption{In this diagram we can appreciate the overall connectivity of the system.
-Note that the inhibitory connections are only fully depicted for the first unit for the sake of clarity.}\label{Fig:diagram}
+\caption{Scheme of the overall network connectivity of the system. Note that the inhibitory connections are only fully depicted for the first unit for the sake of clarity.}\label{Fig:diagram}
 \end{figure}
@@
 This means that we artificially activate the first unit (clamping) of the sequence for a short time.
-After this initial cue, if the weights are set right, the neural network continues evolving on its own and a particular sequence is recalled.
-In figure \ref{Fig:recall} a successful recall process is shown.
+After this initial cue, if the weights have the right configuration, the neural network will continue evolving on its own until the whole particular sequence is correctly recalled as shown in figure  \ref{Fig:recall}.
 
@@
 \includegraphics[scale=0.20]{recall.eps}
-\caption{A succesful recall.
-a) The activity of the units as it evolves in time.
+\caption{Demostration of the successful sequence recall with the corresponding connectivity matrix.
+a) The activity of the units over time.
 b) The currents that each unit receives as the sequence progresses.
@@
 d) The connectivity matrix $W$.
-The weights of the matrix can be divided on the excitatory connections: the ones that induce transitions $Exc_{T}$ and the ones that stabilize the unit $Exc_{self}$ and the inhibitory connections $Inh$.}\label{Fig:recall}
+The weights of the matrix $W$ can be divided into the excitatory connections: the ones on the diagonal that stabilize the unit $Exc_{self}$, the ones under the diagonal that induce transitions $Exc_{T}$ and finally the ones that stabilize the unit  and the inhibitory connections $Inh$ which dominate the matrix $W$.
+e) The legend with the color codes, we use different colors and thickness to represent the different variables corresponding to different units.
+}\label{Fig:recall}
 \end{figure}
@@
 After the first unit $x_1$ has been activated for long enough due to clamping the filter $z_1$ will start catching up with the activity by increasing its value.
-When this happens the unit corresponding to $x_2$ will receive a growing current with value $Exc_{T} \times z_1$.
-As this process continues there will a point in time (exactly at $Exc_{T}(1 - e^{-\frac{t}{\tau_z}}) - \theta = 0$) when $x_2$ will receive enough current to become activated.
-This, will again will prompt the filter $z_2$ to start catching up initializing the process described again.
+It follow then that the unit corresponding to $x_2$ will receive a growing current with value $Exc_{T} \times z_1$ because of the forward connectivity.
+As this process continues there will be a point in time (exactly at $Exc_{T}(1 - e^{-\frac{t}{\tau_z}}) - \theta = 0$) when $x_2$ will receive enough current to become activated.
+This in turn will prompt the filter $z_2$ to start catching up initializing the describe process again.
 On the other hand as soon as the second filter starts growing $z_2$ the first unit will start receiving an inhibitory current $Inh \times z_2$ due to the backward inhibitory projection, this will in turn suppress the unit $x_1$ allowing the sequence to propagate.
@@
 With the knowledge above in mind we can envision the following  failure scenarios for the sequential dynamics.
-First, if the magnitude of $Exc_{self}$ the activity will not be big enough to overcome the threshold and the sequence will not propagate.
-Second, if $Exc_{Self}$ is enough to keep the unit active but $Exc_{T}$ is not big enough the next unit in the sequence will not be activated and the sequence will not propagate.
-If both $Exc_{self}$ is big enough to sustain the activity and $Exc_{T}$ is big enough to propagate it but the inhibitory connection $Inh$ is not enough to suppress the units the result will be that all the units in the sequence will be activated at the end.
+First, if the magnitude of $Exc_{self}$ is small the self-excitatory current will not be big enough to overcome the threshold and the sequence will not propagate.
+Second, if $Exc_{Self}$ is big enough to keep the unit active but $Exc_{T}$ is not big enough the next unit in the sequence will not be activated and the sequence will not propagate, in this case we are stuck in the first pattern forever.
+If both $Exc_{self}$ is big enough to sustain the activity and $Exc_{T}$ is big enough to propagate it but the inhibitory connection $Inh$ is not big enough to suppress the units the result will be that all the units in the sequence will be activated at the end.
 
@@
 
-A common problem in sequence learning is to scale the time duration of the sequence according to the needs of the required context.
-In terms of a neural network it means that once the dynamics of sequence recalling are stored in the shape of the weights we can control the \textit{recall time} of the sequence with the internal parameters of the network where the recall time is the amount of time that a particular unit is active during a sequence recall.
+A common problem in sequence learning is to scale the time duration of the sequence according to the needs of the context.
+In terms of a neural network it means that once the dynamics of sequence recalling are stored in the shape of the weights we can control the \textit{recall time} of the sequence with the internal parameters of the network where we define the recall time as the amount of time that a particular unit is active during a sequence retrieval.
 In our network there are a two parameters that a priori are strong candidates to regulate the recall time: the transition weight $Exc_T$ by controlling how fast you activate the next unit in the sequence and the threshold $\theta$ by controlling how long it takes for the current to activate the next unit in a sequence.
@@
 Once the unit $x_1$ becomes activated the filter $z_1$ starts increasing its value and the unit $x_2$ will be receiving an excitatory current $Exc_T \times z_1$.
-The point in time t at which the this current overcomes the threshold will be the recall time.
+The point in time $t_{recall}$ at which the this current overcomes the threshold will be the recall time.
 Using that reasoning we can solve the following equation to calculate the recall time: $Exc_{T}(1 - e^{-\frac{t}{\tau_z}}) - \theta = 0$.
@@
 \tag{post-synaptic}
-\tau_w \dfrac{dw}{dt} &= \underbrace{(w_{max} - w)}_{\text{saturation}} \underbrace{z_{pre}z_{post}}_{\text{hebbian}} + \underbrace{(w_{min} - w)}_{\text{saturation}}\underbrace{ (1 - z_{pre})z_{post}}_{\text{synaptic-gating}} \label{eq:post-synaptic rule}
+\tau_w \dfrac{dw}{dt} &= \underbrace{(w_{max} - w)}_{\text{saturation}} \underbrace{z_{pre}z_{post}}_{\text{Hebbian}} + \underbrace{(w_{min} - w)}_{\text{saturation}}\underbrace{ (1 - z_{pre})z_{post}}_{\text{synaptic-gating}} \label{eq:post-synaptic rule}
 \end{align}
@@
 \includegraphics[scale=1.3]{plasticity_diagram.eps}
-\caption{A scheme of how plasticity works.
-a) Hebbian term contribution.
-b) The contribution is negative when the post-synaptic unit is active.
-c) The contribution is negative when the pre-synaptic unit is active.
-d) If units are not active there is no decay of the weights.}\label{Fig:plasticity diagram}
+\caption{Conceptual illustration of the mechanism of learning induced synaptic changes.
+Here the a black filling stands for an active unit a) Hebbian associative plasticity that implies a weight increase when both pre- and pos-synaptic units are simultaneously active b) In the case of the post-synaptic rule the weight is decreased if the post-synaptic unit itself active without the pre-synaptic unit being active.
+c) In the case of the pre-synaptic rule the weight is decreased if the pre-synaptic unit is itself active without the post-synaptic unit being active.
+d) If neither pre- nor post-synaptic units are active, there is no weight modification.}\label{Fig:plasticity diagram}
 \end{figure}
@@
 \includegraphics[scale=0.3]{training_rule.eps}
-\caption{Training with the learning rule.
-a) An example of training with a small value of $tau_z$.
-b) Same procedure as in a but with a longer time constant $\tau_z$.
-c) evolution of weights with the number of epochs.
-The three situations depicted utilize the pre-synaptic learning rule.}\label{Fig:epochs}
+\caption{The effect of training with the learning rule.
+a) An example of training with a small value of $tau_z=0.050s$.
+b) Same procedure as in a but with a longer time constant $\tau_z=0.150s$.
+c) evolution in time of the connection weights from the first to the second unit $Exc_T$, the one from the second back to the first unit $Inh$ and finally of the second unit into itself $Exc_{self}$.
+Five epochs of length $2s$ are shown here, note that this is enough time for the weights to stabilize.}\label{Fig:epochs}
 \end{figure}

### 2017-11-29 plot_producers/essan_dynamical_range_double.py

@@ -38,21 +38,24 @@ transition_vector_1 = np.logspace(-14, -0.3, 20) + 0.5
 transition_vector = np.hstack((transition_vector_1, transition_vector_2))
 transition_vector.sort()
 
-recall_times = []
-for transition in transition_vector:
-    w = designed_matrix_sequences(N, sequences, self_excitation=self_excitation, transition=transition,
-                              inhibition=inhibition)
+if True:
+    recall_times = []
+    for transition in transition_vector:
+        w = designed_matrix_sequences(N, sequences, self_excitation=self_excitation, transition=transition,
+                                  inhibition=inhibition)
 
-    dic = run_network_recall_limit(N, w, G, threshold, tau_m, tau_z,  T, dt, I_cue, T_cue)
-    x_history = dic['x']
-    duration = get_recall_duration_for_pattern(x_history, pattern, dt)
-    recall_times.append(duration)
+        dic = run_network_recall_limit(N, w, G, threshold, tau_m, tau_z,  T, dt, I_cue, T_cue)
+        x_history = dic['x']
+        duration = get_recall_duration_for_pattern(x_history, pattern, dt)
+        recall_times.append(duration)
 
-
+else:
+    recall_times =transition_vector
 # The figure
 time_t1_recall = time_t1(tau_z, T=transition_vector, I=inhibition, threshold=threshold)
 
 captions = True
+show_units = True
 
 fig = plt.figure()
 gs = gridspec.GridSpec(2, 2)
@@ -64,8 +67,8 @@ ax2 = fig.add_subplot(gs[1, 1])
 if captions:
     size = 30
     fig.text(0.05, 0.95, 'a)', size=size)
-    fig.text(0.55, 0.95, 'b)', size=size)
-    fig.text(0.55, 0.50, 'c)', size=size)
+    fig.text(0.50, 0.95, 'b)', size=size)
+    fig.text(0.50, 0.50, 'c)', size=size)
 
 plt.tight_layout()
 
@@ -75,7 +78,7 @@ ax.plot(transition_vector, time_t1_recall, label='theoretical')
 ax.axhline(0, ls='--', color='black')
 ax.axvline(threshold, ls=':', color='grey', label=r'threshold $\theta$')
 
-ax.set_title('Dynamical Range')
[…]

### 2017-11-29 plot_producers/essan_recall.py

@@ -109,14 +109,13 @@ ax2.set_xlim([-0.1, 1.8])
 ax3.set_xlim([-0.1, 1.8])
 
 
-if False:
-    ax1.set_xlabel('unit activities, x')
-    ax2.set_ylabel('currents')
-    ax3.set_ylabel('z-filters')
-else:
-    ax1.set_title('unit activities, x')
-    ax2.set_title(r'currents, $\Phi$ argument')
-    ax3.set_title('z-filters')
+ax1.set_ylabel('x')
+ax2.set_ylabel('currents')
+ax3.set_ylabel('z')
+
+ax1.set_title('unit activities')
+ax2.set_title(r'$\Phi$ argument')
+ax3.set_title('z-filters')
 
 ax1.get_xaxis().set_visible(False)
 ax2.get_xaxis().set_visible(False)

### 2017-11-29 plot_producers/essan_training_rule.py

@@ -15,6 +15,7 @@ from network import run_network_recall, train_network, run_network_recall_limit
 
 independent_labels = False  # Whehter you want an extra weight trained matrix or the labels
 captions = True
+pre_rule = False
 
 N = 10
 tau_z = 0.050
@@ -32,7 +33,7 @@ sequence1 = [0, 1, 2, 3, 4, 5, 6, 7, 8, 9]
 sequences = [sequence1]
 
 dic = train_network(N, dt, training_time, inter_sequence_time, sequences, tau_z, tau_z_post,
-                    tau_w, epochs=epochs, max_w=max_w, min_w=min_w, save_w_history=True, pre_rule=True)
+                    tau_w, epochs=epochs, max_w=max_w, min_w=min_w, save_w_history=True, pre_rule=pre_rule)
 
 w = dic['w']
 x_total = dic['x']
@@ -41,7 +42,7 @@ z_post_history = dic['z_post']
 w_history = dic['w_history']
 
 dic = train_network(N, dt, training_time, inter_sequence_time, sequences, tau_z_1, tau_z_post,
-                    tau_w, epochs=epochs, max_w=max_w, min_w=min_w, save_w_history=True, pre_rule=True)
+                    tau_w, epochs=epochs, max_w=max_w, min_w=min_w, save_w_history=True, pre_rule=pre_rule)
 
 w1 = dic['w']
 
@@ -56,9 +57,9 @@ w_11 = w_history[:, 1, 1]
 fig = plt.figure(figsize=(16, 12))
 
 if captions:
-    size = 25
-    fig.text(0.10, 0.85, 'a)', size=size)
-    fig.text(0.55, 0.85, 'b)', size=size)
+    size = 35
+    fig.text(0.10, 0.86, 'a)', size=size)
+    fig.text(0.53, 0.86, 'b)', size=size)
     fig.text(0.10, 0.48, 'c)', size=size)
 
 
@@ -80,8 +81,8 @@ cax = divider.append_axes('right', size='5%', pad=0.05)
 fig.colorbar(im, cax=cax)
 
 ax1.set_title(r'$\tau_z$ = ' + str(tau_z))
-ax1.set_xlabel('Pre')
-ax1.set_ylabel('Post')
+ax1.set_xlabel('pre-synaptic')
[…]

### 2017-11-30 plot_producers/essan_recall.py
Modified the plot producers and added a new notebook

@@ -72,7 +72,7 @@ if captions:
     fig.text(aux_x, 0.60, 'b)', size=size)
     fig.text(aux_x, 0.35, 'c)', size=size)
     fig.text(0.5, 0.90, 'd)', size=size)
-    fig.text(0.5, 0.40, 'e)', size=size)
+    # fig.text(0.5, 0.40, 'e)', size=size)
 
 
 ax1 = fig.add_subplot(gs[0, 0])

### 2017-11-30 plot_producers/essan_training_rule_quantities.py

@@ -17,8 +17,33 @@ from network import train_network, run_network_recall_limit
 legends_instead_of_post = True
 add_max_and_min_markers = True
 captions = True
-pre_rule = True
+pre_rule = False
+invert = True
 
+
+N = 10
+inter_sequence_time = 1.000
+dt = 0.001
+sequence = [0, 1, 2, 3, 4, 5, 6, 7, 8, 9]
+sequences = [sequence]
+
+if pre_rule:
+    tau_z_post_general = 0.005
+    tau_w_general = 0.100
+    training_time_general = 0.500
+    tau_z_general = 0.050
+    max_w_general = 20.0
+    min_w_general = -10.0
+    epochs_general = 4
+
+else:
+    tau_z_post_general = 0.005
+    tau_w_general = 0.050
+    training_time_general = 0.500
+    tau_z_general = 0.100
+    max_w_general = 5.0
+    min_w_general = -10.0
+    epochs_general = 4
 
 fig = plt.figure(figsize=(16, 12))
 ax1 = fig.add_subplot(321)
@@ -30,27 +55,21 @@ ax5 = fig.add_subplot(325)
 ax6 = fig.add_subplot(326)
 
 if captions:
-    size = 25
+    size = 30
     fig.text(0.05, 0.97, 'a)', size=size)
     fig.text(0.50, 0.97, 'b)', size=size)
     fig.text(0.05, 0.65, 'c)', size=size)
-    fig.text(0.05, 0.32, 'd)', size=size)
+    fig.text(0.05, 0.34, 'd)', size=size)
     fig.text(0.50, 0.32, 'f)', size=size)
 
 
 plt.tight_layout()
 
 # Some general parameters
-N = 10
-inter_sequence_time = 1.000
-dt = 0.001
-sequence1 = [0, 1, 2, 3, 4, 5, 6, 7, 8, 9]
-sequences = [sequence1]
-epochs = 5
+epochs = epochs_general
+
 pattern = 3
 pattern_from = 2
-
-
 
 # Training time
 training_times_vector = np.arange(0.050, 1.050, 0.050)
@@ -58,13 +77,15 @@ w_self = []
 w_transition = []
 w_inhition = []
 
-tau_z = 0.050
-tau_z_post = 0.005
-tau_w = 0.100
-max_w = 10.0
-min_w = -3.0
+tau_z = tau_z_general
+tau_z_post = tau_z_post_general
+tau_w = tau_w_general
+max_w = max_w_general
[…]

### 2017-11-30 ESSAN/article_before_triming.tex
Uploaded the article

\documentclass{esannV2}
\usepackage[dvips]{graphicx}
\usepackage[latin1]{inputenc}
\usepackage{float}
\usepackage{subcaption}
\usepackage{amssymb,amsmath,array}
\graphicspath{{images/}}

%***********************************************************************
% !!!! IMPORTANT NOTICE ON TEXT MARGINS !!!!!
%***********************************************************************
%
% Please avoid using DVI2PDF or PS2PDF converters: some undesired
% shifting/scaling may occur when using these programs
% It is strongly recommended to use the DVIPS converters, and to submit
% PS file. You may submit a PDF file if and only if you use ADOBE ACROBAT
% to convert your PS file to PDF.
%
% Check that you have set the paper size to A4 (and NOT to letter) in your
% dvi2ps converter, in Adobe Acrobat if you use it, and in any printer driver
% that you could use.  You also have to disable the 'scale to fit paper' option
% of your printer driver.
%
% In any case, please check carefully that the final size of the top and
% bottom margins is 5.2 cm and of the left and right margins is 4.4 cm.
% It is your responsibility to verify this important requirement.  If these margin requirements and not fulfilled at the end of your file generation process, please use the following commands to correct them.  Otherwise, please do not modify these commands.
%
\voffset 0 cm \hoffset 0 cm \addtolength{\textwidth}{0cm}
\addtolength{\textheight}{0cm}\addtolength{\leftmargin}{0cm}

%***********************************************************************
% !!!! USE OF THE esannV2 LaTeX STYLE FILE !!!!!
%***********************************************************************
%
% Some commands are inserted in the following .tex example file.  Therefore to
% set up your ESANN submission, please use this file and modify it to insert
% your text, rather than staring from a blank .tex file.  In this way, you will
% have the commands inserted in the right place.

\begin{document}
%style file for ESANN manuscripts
\title{Sequence learning and recall in recurrent neural networks with time filtered units.}

%***********************************************************************
% AUTHORS INFORMATION AREA
%***********************************************************************
\author{Ram\'on H. Mart\'inez-Mayorquin $^1$, Anders Lansner and $^1$ Pawel Herman $^1$
%
% Optional short acknowledgment: remove next line if non-needed
% \thanks{This is an optional funding source acknowledgement.}
%
% DO NOT MODIFY THE FOLLOWING '\vspace' ARGUMENT
\vspace{.3cm}\\
%
% Addresses and institutions (remove "1- " in case of a single institution)
1- KTH Royal Institute of Technology \\
Department of Computational Science and Technology - Stockholm, Sweden.
%
% Remove the next three lines in case of a single institution
\vspace{.1cm}\\ 
}

%***********************************************************************
% END OF AUTHORS INFORMATION AREA
%***********************************************************************

\maketitle

\begin{abstract}
We demonstrate among others how the learned sequences can be recalled at different speeds largely independent of the encoding process. diojafoids sf dsfsd siofjsdoif
sifjsodif sdjfiosdjf sdio jsdoi fjsoijgi jsoifd jdoifg jodifj iodfj odifjg idfjg
fdjgidfjg djf oidfjgiodfj jdf idfjg odfijg difjg odifjg difg jj dijdfgoidf 
jdfiogjdiofgj 
\end{abstract}

\section{Introduction}
There is a growing interest in studying the capability to learn, store and flexibly recall sequential patterns in neural network models. Processing sequential information bears particular relevance to a wide scope of cognitive functionality and behavior in humans \cite{lashley1951problem}. Hebb proposed that sequential activation of neural cell assemblies, referred to as phases sequences, form the basis of the human thought process \cite{hebb2005organization}. Serial order can be attributed to speech perception and generation, motor control and planning as well as goal directed behavior orchestrated by working memory. Therefore the question as to how a novel sequence of items is stored and recalled in the corrected order has been extensively studied in the context of memory systems among others. Besides intensive efforts in this direction undertaken in the domain of cognitive neuroscience and psychology \cite{hurlstone2014memory} the problem of sequence learning has also attracted a lot of theoretical interest. It has resulted in the development of different computational network models simulating the brain's capabilities to process sequence information. A prominent share of this computaitonal endevaour is linked to spiking neural network approaches to learn precise patterns of spike sequences with biologically plausible local learning rules \cite{ans1994neural}\cite{dehaene1987neural}\cite{gutig2006tempotron}\cite{ponulak2010supervised}. Another group of network approaches developed in the realm of rate based models offers a more generic context for sequence memory function \cite{amari1972learning}\cite{kohonen1977principle}\cite{sandberg2002bayesian}. The work presented here belongs to that category of studies and conceptually builds on \cite{sandberg2002bayesian}. In particular we rely on a simple recurrent neural network architecture and incorporate the concept of synaptic filtering into a basic Hebbian learning rule with synaptic gating \cite{andrew2003spiking} to study the network's capabilities to learn and recall sequences of memory items. The proposed rate based model constitutes a well-controlled framework for studying the fundamental synaptic mechanisms and identifying key parameters responsible for sequence processing capabilities in a dynamical connectionist system with all-to-all recurrent connectivity pattern.

\section{Methods}

The recurrent neural network in Fig. \ref{Fig:diagrams} is composed of firing rate units with the input-output characteristics described by the logistic function $\Phi(x) = \frac{1}{1 + \exp(-Gx)}$ with gain G. The synaptic inputs are aggregated across the entire network through all-to-all connections weighted by synaptic weights $W$, and the activation of unit $U_i$, is then consistently determined by integrating synaptic inputs with the time constant $\tau_{m}$ (\ref{eq1}). To account for the temporal effect of synaptic smoothing we introduce the idea of z-filters  \cite{tully2016spike} that maintain the exponentially decaying traces of the activity with the time constant of $\tau_z$ (\ref{eq2}). This allows for keeping synaptic memory of the unit activity and influencing other post-synaptic units in the network even when the original pre-synaptic unit is not longer active.

\begin{align}
\tau_m \dfrac{dx_i}{dt} &= \Phi\Big(\sum_{j} W_{ij} z_j - \theta \Big) - x_i \label{eq1} \\ 
\tau_z \dfrac{dz_i}{dt} &= x_i - z_i \label{eq2}
\end{align}

In this work we examine two types of learning rules, pre-synaptic (3) and post-synaptic (4)
gating \cite{andrew2003spiking}, which equip the proposed network with the capability to learn
how to recall sequential activity patterns.

\begin{align}
\tag{3 pre-synaptic}
\tau_w \dfrac{dw}{dt} &= (w_{max} - w) z_{pre}z_{post} + (w_{min} - w) (1 - z_{post})z_{pre} \label{eq:pre-synaptic rule}  \\
\tag{4 post-synaptic}
\tau_w \dfrac{dw}{dt} &= \underbrace{(w_{max} - w)}_{\text{saturation}} \underbrace{z_{pre}z_{post}}_{\text{Hebbian}} + \underbrace{(w_{min} - w)}_{\text{saturation}}\underbrace{ (1 - z_{pre})z_{post}}_{\text{synaptic-gating}} \label{eq:post-synaptic rule}
\end{align}

\begin{figure}[H]
\centering
\includegraphics[scale=0.8]{diagram_mixed.eps}
\caption{a) Scheme of the overall network connectivity of the system. Note that the inhibitory connections are only fully depicted for the first unit for the sake of clarity. b) Conceptual illustration of the mechanism of learning induced synaptic changes. Here the a black filling stands for an active unit. } \label{Fig:diagrams}
\end{figure}

The first term in (\ref{eq:pre-synaptic rule}) and (\ref{eq:post-synaptic rule}) is an associative Hebbian contribution \cite{hebb2005organization} and its role is to strengthen the connections between activated units. Importantly however, in comparison with the original formulation \cite{andrew2003spiking}, we use the z-filters instead of the unit activities. This provides the scope for making associative connections between units even when either of the two units is no longer active provided that its memory trace is retainedby the synapse. As mentioned earlier, the synaptic trace decays exponentially with the time constant of $\tau_z$ (\ref{eq2}), and to preserve serial recall order we introduce asymmetrical time constants for z-pre and z-post : $\tau_{z_{pre}} > \tau_{z_{post}}$. The second term is called synaptic gating \cite{andrew2003spiking}, which facilitates the formation of inhibitory connections whenever the activity (or its z-filtered trace) of one unit is not accompanied with the activity of the other unit. We examine two variations of this term to explore the functional implications of both pre- and post-synaptic gating. Finally, the weights are kept between $w_{min}$ and $w_{max}$ limits. The nature of synaptic changes under the synaptic gating rule is illustrated in Fig. \ref{Fig:diagrams}.

%\begin{figure}[h!]
%\centering
%\includegraphics[scale=1.3]{plasticity_diagram.eps}
%\caption{Conceptual illustration of the mechanism of learning induced synaptic changes. Here the a black filling stands for an active unit a) Hebbian associative plasticity that implies a weight increase when both pre- and pos-synaptic units are simultaneously active b) In the case of the post-synaptic rule the weight is decreased if the post-synaptic unit itself active without the pre-synaptic unit being active. c) In the case of the pre-synaptic rule the weight is decreased if the pre-synaptic unit is itself active without the post-synaptic unit being active. d) If neither pre- nor post-synaptic units are active, there is no weight modification.}\label{Fig:plasticity diagram}
%\end{figure}
%
As a result of learning described above, three types of connections emerge from each unit (Fig. \ref{Fig:diagrams}): two excitatory connections - self-reinforcing unit activity, $Exc_{self}$, and exciting the following unit in the sequence (contributing to the sequential structure of the activity) $Exc_T$, and a range of inhibitory connections, $Inh$, to all the remaining units (among inhibitory connections of special relevance are the backwards inhibitory projections that suppress the unit activity once the next unit in the sequence gets activated).

\section{Results}

\subsection{The neural network can sustain sequential activity}

To recall a particular sequence we cue the first sequence item (unit) by injecting a short input lasting around $\tau_z$. After this initial cue, if the weights have the right configuration as a result of training, the neural network continues to visit subsequent items in the learned sequence to the very
end, as shown in Fig. \ref{Fig:recall}.

\begin{figure}[h!]
\centering
\includegraphics[scale=0.20]{recall.eps}
\caption{Demostration of the successful sequence recall with the corresponding connectivity matrix. a) The activity of the units over time. b) The currents that each unit receives as the sequence progresses. c) the z-filters activations for each unit. d) The connectivity matrix $W$. The weights of the matrix $W$ can be divided into the excitatory connections: the ones on the diagonal that stabilize the unit $Exc_{self}$, the ones under the diagonal that induce transitions $Exc_{T}$ and finally the ones that stabilize the unit  and the inhibitory connections $Inh$ which dominate the matrix $W$. We use different colors to represent different items. }\label{Fig:recall}
\end{figure}

The dynamics of the sequential activity propagation can be easily explained. After the first unit $U_1$ has been activated by the external input, the corresponding filter $z_1$ starts following the activity $x_i$. In consequence, the subsequent unit $U_2$ receives a growing current with value $z_1 Exc_T$ because of the forward connectivity. This process continues until the point in time when $Exc_{T}(1 - e^{-\frac{t}{\tau_z}}) - \theta = 0$ implying that $U_2$ receives enough current to become activated. This in turn prompts the filter $z_2$ to follow the activity $x_2$ initializing again the process described above. On the other hand as soon as the synaptic trace corresponding to $x_2$ starts growing, the preceding unit $U_1$ beings to receive an inhibitory current $z_2 Inh$ due to the backward inhibitory projection. This in turn suppresses the activity of unit $U_1$ facilitating the propagation of the sequence. Finally once the unit $U_1$ is completely suppressed the excitatory current mediated by $Exc_T$ is diminished, which in most cases would also suppress the activity $x_2$ if it was not for the self-sustaining current $z_2 Exc_{self}$. In short, the term $Exc_T$ ensures the transition of the activity from one network state (sequence item) to another one (the next item in the sequence) whereas the competition between $Exc_{self}$ ad $Inh$ first lock the pattern unit and then finally suppresses it enabling the sequence to continue its trajectory.

With the knowledge above in mind we can envision the following  failure scenarios for the sequential dynamics. First, if the magnitude of $Exc_{self}$ weight is small the self-excitatory current is not sufficiently strong to overcome the threshold and the sequence stalls. Second, if $Exc_{self}$ is large enough to keep the unit active but the $Exc_{T}$ weight is weak, the next unit in the sequence does not become activated and the sequence remains stuck forever in the same state corresponding to the first pattern. Finally if both $Exc_{self}$ and $Exc_T$ are strong enough to sustain and propagate the activity respectively but the inhibitory weight $Inh$ is not sufficiently large to suppress the activity, all the units in the sequence become co-activated at the end.

\subsection{The neural network possess a wide dynamic range}

A common problem in sequence learning is to scale the time duration of the sequence according to the contextual demands \cite{murray2017learning}. For a neural network it means that the dynamics of sequence recall can be flexibly controlled with the network parameters after the training process is completed. The recall time can be defined as the temporal length of the activation of the individual units during the sequence retrieval. In the network there are two key parameters that are directly implicated in the regulation of the recall time: the transition weight, $Exc_T$ and the threshold, $\theta$, which control how long it takes to recruit the next unit in a sequence. Below we derive the estimate of how $Exc_T$ affects the recall time and the reasoning for $\theta$ is analogous.

In order to calculate the recall time we make the following consideration. Once the unit $U_1$ becomes activated the filter $z_1$ starts growing thereby contribution to the excitatory current input, $z_1 Exc_T$, to the next unit $U_2$. The point in time at which this current overcomes the threshold marks the end of the recall process for a specific item in the sequence. Following that reasoning we can solve the equation $Exc_{T}(1 - e^{-\frac{t}{\tau_z}}) - \theta = 0$ to estimate the recall time. As a result, we obtain the following expression:

\begin{align}
T_{recall} = \tau_z \log\Big(\frac{Exc_T}{Exc_T - \theta} \Big)
\end{align}

This expression has a singularity at $Exc_T = \theta$, which implies that the recall time can be made arbitrarily long by setting $Exc_T$ close to the threshold. However due to the all-to-all connectivity the are other contributions which we can take into account with:

\begin{align}
T_{recall} = \tau_z \log\Big(\frac{Exc_T + Inh}{Exc_T - \theta}\Big) \label{eq:recall_complex}
\end{align}

Figure \ref{Fig:dynamical_range} illustrates the wide dynamical range of the recall time in our network model (both simulated and estimated as presented above)

\begin{figure}[h!]
\centering
\includegraphics[scale=0.28]{dynamical_range.eps}
\caption{The network posses a wide dynamical range. a) Here we compare the calculated thoretical expression with simulations with our network for different values of $Exc_T$. Note the singularity at $\theta$. b) The network activity with a short recall time. c) The network activity with a longer recall time ($Exc_T$ close to $\theta$).}\label{Fig:dynamical_range}
\end{figure}

\subsection{Learning Rule}
Our learning protocol consists in fixing (clamping) each unit of the sequence in succession for a given amount of time (training time). We perform this process an epochs number of times. It is important to note that while learning there is a period of one second between each epoch or presentation that serves the purposes of avoiding self-interference.

\begin{figure}[H]
\centering
\includegraphics[scale=0.3]{training_rule.eps}
\caption{The effect of training with the learning rule. a) An example of training with a small value of $tau_z=0.050s$. b) Same procedure as in a but with a longer time constant $\tau_z=0.150s$. c) evolution in time of the connection weights from the first to the second unit $Exc_T$, the one from the second back to the first unit $Inh$ and finally of the second unit into itself $Exc_{self}$. Five epochs of length $2s$ are shown here, note that this is enough time for the weights to stabilize.}\label{Fig:epochs}
\end{figure}

%\begin{figure}[H]
%\centering
%\includegraphics[scale=0.3]{training_rule_quantities_pre_rule_True.eps}
%\caption{Learning dynamics of the pre-synaptic learning rule.}\label{Fig:learning_quantities_pre}
%\end{figure}
%
%
%\textbf{Training time}
%\begin{itemize}
%\item  $Exc_{self}$: The longer the training time the longer the filter is activated with itself and therefore it grows. But given that the pre and post filters do not coincide completely in the temporal domain there are negative contributions which prevent this from reaching the saturation value $w_{max}$
%\item  $Inh$: In this case the the negative contributions come from the time that the units are not activated together with their counterparts which is most of the time (except in the brief transition period) and therefore the value saturates quickly to $w_{min}$.
%\item $Exc_{T}$: In this case what matters is the transition time between two 
%\end{itemize} 
%
%
%\begin{figure}[H]
%\centering
%\includegraphics[scale=0.3]{training_rule_quantities_pre_rule_False.eps}
%\caption{Learning dynamics of the post-synaptic learning rule}\label{Fig:learning_quantities_post}
%\end{figure}

\section{Discussion}

% ****************************************************************************
% BIBLIOGRAPHY AREA
% ****************************************************************************

\begin{footnotesize}

% IF YOU USE BIBTEX,
% - DELETE THE TEXT BETWEEN THE TWO ABOVE DASHED LINES
% - UNCOMMENT THE NEXT TWO LINES AND REPLACE 'Name_Of_Your_BibFile'

\bibliographystyle{unsrt}
\bibliography{references.bib}

\end{footnotesize}

% ****************************************************************************
% END OF BIBLIOGRAPHY AREA
% ****************************************************************************

\end{document}

### 2018-09-10 ESSAN/swedcog.tex
Added disambiguation case

--- 2017-11-30 ESSAN/article_before_triming.tex
@@
-\documentclass{esannV2}
-\usepackage[dvips]{graphicx}
-\usepackage[latin1]{inputenc}
+\documentclass[10pt, a4paper]{article}
+\usepackage[left=2cm, right=2cm, top=2cm]{geometry}
+\usepackage[utf8]{inputenc}
+\usepackage{amsmath}
+\usepackage{amsfonts}
+\usepackage{amssymb}
+\usepackage{amsthm}
+\usepackage{mathptmx}
+
 \usepackage{float}
-\usepackage{subcaption}
-\usepackage{amssymb,amsmath,array}
+\usepackage{subfigure}
+\usepackage{framed}
+\usepackage{graphicx}
 \graphicspath{{images/}}
 
-%***********************************************************************
-% !!!! IMPORTANT NOTICE ON TEXT MARGINS !!!!!
-%***********************************************************************
-%
-% Please avoid using DVI2PDF or PS2PDF converters: some undesired
-% shifting/scaling may occur when using these programs
-% It is strongly recommended to use the DVIPS converters, and to submit
-% PS file. You may submit a PDF file if and only if you use ADOBE ACROBAT
-% to convert your PS file to PDF.
-%
-% Check that you have set the paper size to A4 (and NOT to letter) in your
-% dvi2ps converter, in Adobe Acrobat if you use it, and in any printer driver
-% that you could use.  You also have to disable the 'scale to fit paper' option
-% of your printer driver.
-%
-% In any case, please check carefully that the final size of the top and
-% bottom margins is 5.2 cm and of the left and right margins is 4.4 cm.
-% It is your responsibility to verify this important requirement.
-If these margin requirements and not fulfilled at the end of your file generation process, please use the following commands to correct them.
-Otherwise, please do not modify these commands.
-%
-\voffset 0 cm \hoffset 0 cm \addtolength{\textwidth}{0cm}
-\addtolength{\textheight}{0cm}\addtolength{\leftmargin}{0cm}
+% APA citation style
+\usepackage{apacite}
 
-%***********************************************************************
-% !!!! USE OF THE esannV2 LaTeX STYLE FILE !!!!!
-%***********************************************************************
-%
-% Some commands are inserted in the following .tex example file.  Therefore to
-% set up your ESANN submission, please use this file and modify it to insert
-% your text, rather than staring from a blank .tex file.  In this way, you will
-% have the commands inserted in the right place.
-
-\begin{document}
-%style file for ESANN manuscripts
 \title{Sequence learning and recall in recurrent neural networks with time filtered units.}
-
-%***********************************************************************
-% AUTHORS INFORMATION AREA
-%***********************************************************************
 \author{Ram\'on H. Mart\'inez-Mayorquin $^1$, Anders Lansner and $^1$ Pawel Herman $^1$
@@
 % Remove the next three lines in case of a single institution
-\vspace{.1cm}\\ 
+\vspace{.3cm}\\ 
 }
 
-%***********************************************************************
-% END OF AUTHORS INFORMATION AREA
-%***********************************************************************
-
+\setlength{\parskip}{0.3cm}
+\begin{document}
 \maketitle
 
-\begin{abstract}
-We demonstrate among others how the learned sequences can be recalled at different speeds largely independent of the encoding process. diojafoids sf dsfsd siofjsdoif
-sifjsodif sdjfiosdjf sdio jsdoi fjsoijgi jsoifd jdoifg jodifj iodfj odifjg idfjg
-fdjgidfjg djf oidfjgiodfj jdf idfjg odfijg difjg odifjg difg jj dijdfgoidf 
-jdfiogjdiofgj 
-\end{abstract}
-
-\section{Introduction}
 There is a growing interest in studying the capability to learn, store and flexibly recall sequential patterns in neural network models.
@@
 Another group of network approaches developed in the realm of rate based models offers a more generic context for sequence memory function \cite{amari1972learning}\cite{kohonen1977principle}\cite{sandberg2002bayesian}.
-The work presented here belongs to that category of studies and conceptually builds on \cite{sandberg2002bayesian}.
-In particular we rely on a simple recurrent neural network architecture and incorporate the concept of synaptic filtering into a basic Hebbian learning rule with synaptic gating \cite{andrew2003spiking} to study the network's capabilities to learn and recall sequences of memory items.
-The proposed rate based model constitutes a well-controlled framework for studying the fundamental synaptic mechanisms and identifying key parameters responsible for sequence processing capabilities in a dynamical connectionist system with all-to-all recurrent connectivity pattern.
 
-\section{Methods}
+The work presented here belongs to that category of connectionist based rate models and conceptually builds on \cite{sandberg2002bayesian}.
+As our main contribution we propose the use of a new Hebbian learning associative rule with synaptic gating \cite{andrew2003spiking} that provide a simple recurrent neural network with capabilities to learn and recall sequences of memory items.
+The proposed rate based model constitutes a well-controlled framework for studying the both learning and processing of sequences at the sub-symbolic level and identifying key parameters responsible for sequence processing capabilities in a dynamical connectionist system with all-to-all recurrent connectivity pattern.
 
@@
 
-%\begin{figure}[h!]
-%\centering
-%\includegraphics[scale=1.3]{plasticity_diagram.eps}
-%\caption{Conceptual illustration of the mechanism of learning induced synaptic changes.
-Here the a black filling stands for an active unit a) Hebbian associative plasticity that implies a weight increase when both pre- and pos-synaptic units are simultaneously active b) In the case of the post-synaptic rule the weight is decreased if the post-synaptic unit itself active without the pre-synaptic unit being active.
-c) In the case of the pre-synaptic rule the weight is decreased if the pre-synaptic unit is itself active without the post-synaptic unit being active.
-d) If neither pre- nor post-synaptic units are active, there is no weight modification.}\label{Fig:plasticity diagram}
-%\end{figure}
-%
-As a result of learning described above, three types of connections emerge from each unit (Fig.
-\ref{Fig:diagrams}): two excitatory connections - self-reinforcing unit activity, $Exc_{self}$, and exciting the following unit in the sequence (contributing to the sequential structure of the activity) $Exc_T$, and a range of inhibitory connections, $Inh$, to all the remaining units (among inhibitory connections of special relevance are the backwards inhibitory projections that suppress the unit activity once the next unit in the sequence gets activated).
-
-\section{Results}
-
-\subsection{The neural network can sustain sequential activity}
-
-To recall a particular sequence we cue the first sequence item (unit) by injecting a short input lasting around $\tau_z$.
-After this initial cue, if the weights have the right configuration as a result of training, the neural network continues to visit subsequent items in the learned sequence to the very
-end, as shown in Fig. \ref{Fig:recall}.
-
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=0.20]{recall.eps}
-\caption{Demostration of the successful sequence recall with the corresponding connectivity matrix.
-a) The activity of the units over time.
-b) The currents that each unit receives as the sequence progresses.
-c) the z-filters activations for each unit.
-d) The connectivity matrix $W$.
-The weights of the matrix $W$ can be divided into the excitatory connections: the ones on the diagonal that stabilize the unit $Exc_{self}$, the ones under the diagonal that induce transitions $Exc_{T}$ and finally the ones that stabilize the unit  and the inhibitory connections $Inh$ which dominate the matrix $W$.
-We use different colors to represent different items.
-}\label{Fig:recall}
-\end{figure}
-
-The dynamics of the sequential activity propagation can be easily explained.
-After the first unit $U_1$ has been activated by the external input, the corresponding filter $z_1$ starts following the activity $x_i$.
-In consequence, the subsequent unit $U_2$ receives a growing current with value $z_1 Exc_T$ because of the forward connectivity.
-This process continues until the point in time when $Exc_{T}(1 - e^{-\frac{t}{\tau_z}}) - \theta = 0$ implying that $U_2$ receives enough current to become activated.
-This in turn prompts the filter $z_2$ to follow the activity $x_2$ initializing again the process described above.
-On the other hand as soon as the synaptic trace corresponding to $x_2$ starts growing, the preceding unit $U_1$ beings to receive an inhibitory current $z_2 Inh$ due to the backward inhibitory projection.
-This in turn suppresses the activity of unit $U_1$ facilitating the propagation of the sequence.
-Finally once the unit $U_1$ is completely suppressed the excitatory current mediated by $Exc_T$ is diminished, which in most cases would also suppress the activity $x_2$ if it was not for the self-sustaining current $z_2 Exc_{self}$.
-In short, the term $Exc_T$ ensures the transition of the activity from one network state (sequence item) to another one (the next item in the sequence) whereas the competition between $Exc_{self}$ ad $Inh$ first lock the pattern unit and then finally suppresses it enabling the sequence to continue its trajectory.
-
-With the knowledge above in mind we can envision the following  failure scenarios for the sequential dynamics.
-First, if the magnitude of $Exc_{self}$ weight is small the self-excitatory current is not sufficiently strong to overcome the threshold and the sequence stalls.
-Second, if $Exc_{self}$ is large enough to keep the unit active but the $Exc_{T}$ weight is weak, the next unit in the sequence does not become activated and the sequence remains stuck forever in the same state corresponding to the first pattern.
-Finally if both $Exc_{self}$ and $Exc_T$ are strong enough to sustain and propagate the activity respectively but the inhibitory weight $Inh$ is not sufficiently large to suppress the activity, all the units in the sequence become co-activated at the end.
-
-\subsection{The neural network possess a wide dynamic range}
-
-A common problem in sequence learning is to scale the time duration of the sequence according to the contextual demands \cite{murray2017learning}.
-For a neural network it means that the dynamics of sequence recall can be flexibly controlled with the network parameters after the training process is completed.
-The recall time can be defined as the temporal length of the activation of the individual units during the sequence retrieval.
-In the network there are two key parameters that are directly implicated in the regulation of the recall time: the transition weight, $Exc_T$ and the threshold, $\theta$, which control how long it takes to recruit the next unit in a sequence.
-Below we derive the estimate of how $Exc_T$ affects the recall time and the reasoning for $\theta$ is analogous.
-
-In order to calculate the recall time we make the following consideration.
-Once the unit $U_1$ becomes activated the filter $z_1$ starts growing thereby contribution to the excitatory current input, $z_1 Exc_T$, to the next unit $U_2$.
-The point in time at which this current overcomes the threshold marks the end of the recall process for a specific item in the sequence.
-Following that reasoning we can solve the equation $Exc_{T}(1 - e^{-\frac{t}{\tau_z}}) - \theta = 0$ to estimate the recall time.
-As a result, we obtain the following expression:
-
-\begin{align}
-T_{recall} = \tau_z \log\Big(\frac{Exc_T}{Exc_T - \theta} \Big)
-\end{align}
-
-This expression has a singularity at $Exc_T = \theta$, which implies that the recall time can be made arbitrarily long by setting $Exc_T$ close to the threshold.
-However due to the all-to-all connectivity the are other contributions which we can take into account with:
-
-\begin{align}
-T_{recall} = \tau_z \log\Big(\frac{Exc_T + Inh}{Exc_T - \theta}\Big) \label{eq:recall_complex}
-\end{align}
-
-Figure \ref{Fig:dynamical_range} illustrates the wide dynamical range of the recall time in our network model (both simulated and estimated as presented above)
-
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=0.28]{dynamical_range.eps}
-\caption{The network posses a wide dynamical range.
-a) Here we compare the calculated thoretical expression with simulations with our network for different values of $Exc_T$.
-Note the singularity at $\theta$.
-b) The network activity with a short recall time.
-c) The network activity with a longer recall time ($Exc_T$ close to $\theta$).}\label{Fig:dynamical_range}
-\end{figure}
-
-\subsection{Learning Rule}
 Our learning protocol consists in fixing (clamping) each unit of the sequence in succession for a given amount of time (training time).
@@
 \caption{The effect of training with the learning rule.
-a) An example of training with a small value of $tau_z=0.050s$.
+a) An example of training with a small value of $\tau_z=0.050s$.
 b) Same procedure as in a but with a longer time constant $\tau_z=0.150s$.
@@
 
-%\begin{figure}[H]
-%\centering
-%\includegraphics[scale=0.3]{training_rule_quantities_pre_rule_True.eps}
-%\caption{Learning dynamics of the pre-synaptic learning rule.}\label{Fig:learning_quantities_pre}
-%\end{figure}
-%
-%
-%\textbf{Training time}
-%\begin{itemize}
-%\item  $Exc_{self}$: The longer the training time the longer the filter is activated with itself and therefore it grows.
-But given that the pre and post filters do not coincide completely in the temporal domain there are negative contributions which prevent this from reaching the saturation value $w_{max}$
-%\item  $Inh$: In this case the the negative contributions come from the time that the units are not activated together with their counterparts which is most of the time (except in the brief transition period) and therefore the value saturates quickly to $w_{min}$.
-%\item $Exc_{T}$: In this case what matters is the transition time between two 
-%\end{itemize} 
-%
-%
-%\begin{figure}[H]
-%\centering
-%\includegraphics[scale=0.3]{training_rule_quantities_pre_rule_False.eps}
-%\caption{Learning dynamics of the post-synaptic learning rule}\label{Fig:learning_quantities_post}
-%\end{figure}
-
-\section{Discussion}
-
-% ****************************************************************************
-% BIBLIOGRAPHY AREA
-% ****************************************************************************
-
-\begin{footnotesize}
-
-% IF YOU USE BIBTEX,
-% - DELETE THE TEXT BETWEEN THE TWO ABOVE DASHED LINES
-% - UNCOMMENT THE NEXT TWO LINES AND REPLACE 'Name_Of_Your_BibFile'
-
-\bibliographystyle{unsrt}
+%\bibliographystyle{apacite}
+\bibliographystyle{apacitex}
+%\bibliographystyle{apacann}
+%\bibliographystyle{apacannx}
+%\bibliographystyle{apacite}
 \bibliography{references.bib}
-
-\end{footnotesize}
-
-% ****************************************************************************
-% END OF BIBLIOGRAPHY AREA
-% ****************************************************************************
-
 \end{document}

### 2018-09-10 analysis.py

@@ -12,11 +12,11 @@ def get_recall_duration_for_pattern(x_history, pattern, dt):
     return duration
 
 
-def get_recall_duration_sequence(x_history):
+def get_recall_duration_sequence(x_history, dt):
 
     durations = []
     for pattern in range(x_history.shape[1]):
-        duration = get_recall_duration_for_pattern(x_history, pattern)
+        duration = get_recall_duration_for_pattern(x_history, pattern, dt)
         durations.append(duration)
 
     return np.nanmedian(durations[1:-1]), np.nanstd(durations[1:-1]), durations

### 2018-09-10 network.py

@@ -42,10 +42,15 @@ def run_network_recall(N, w, G, threshold, tau_m, tau_z,  T, dt, I_cue, T_cue, s
 
 
 def train_network(N, dt, training_time, inter_sequence_time, sequences, tau_z, tau_z_post, tau_w,
-                  epochs=1, max_w=1.0, min_w=None, save_w_history=False, pre_rule=True):
+                  epochs=1, max_w=1.0, min_w=None, gh=1.0, gah=1.0, save_w_history=False, pre_rule=True):
 
     w = np.zeros((N, N))
     w_history = [w]
+    normal_history = []
+    negative_history = []
+    increase_history = []
+    decrease_history = []
+
 
     inter_sequence_steps = int(inter_sequence_time / dt)
 
@@ -84,28 +89,39 @@ def train_network(N, dt, training_time, inter_sequence_time, sequences, tau_z, t
         if pre_rule:
             negative = np.outer(1 - z_post, z)
         else:
-            negative = np.outer(1 - z, z_post)
+            negative = np.outer(z_post, 1 - z)
+
         if min_w is None:
             w += (dt / tau_w) * ((max_w - w) * normal - negative)
         else:
-            w += (dt / tau_w) * ((max_w - w) * normal + (min_w - w) * negative)
+            increase_terms = (max_w - w) * normal
+            decrease_terms = (min_w - w) * negative
+            w += (dt / tau_w) * (gh * increase_terms + gah * decrease_terms)
 
         if save_w_history:
             w_history.append(np.copy(w))
+            negative_history.append(np.copy(negative))
+            normal_history.append(np.copy(normal))
+            decrease_history.append(np.copy(decrease_terms))
+            increase_history.append(np.copy(increase_terms))
 
     dic = {}
     dic['w'] = w
-    dic['x'] = x_example
+    dic['x'] = x_total
     dic['z'] = z_history
     dic['z_post'] = z_post_history
 
     if save_w_history:
         dic['w_history'] = np.array(w_history)
+        dic['normal'] = np.array(normal_history)
+        dic['negative'] = np.array(negative_history)
+        dic['decrease history'] = np.array(decrease_history)
+        dic['increase history'] = np.array(increase_history)
 
     return dic
 
 
[…]

### 2019-01-18 ESSAN/article.tex
last batch updated

--- 2017-11-29 ESSAN/article.tex
@@
-\documentclass{esannV2}
+\documentclass[a4paper]{esannV2}
 \usepackage[dvips]{graphicx}
 \usepackage[latin1]{inputenc}
+\usepackage{float}
+\usepackage{subcaption}
 \usepackage{amssymb,amsmath,array}
@@
 %style file for ESANN manuscripts
-\title{LaTeX style file for ESANN manuscripts}
+\title{Sequence learning and recall in recurrent neural networks with memory traces}
 
@@
 %***********************************************************************
-\author{Ram\'on H. Mart\'inez-Mayorquin $^1$, Anders Lansner and $^2$ Pawel Herman $^3$
+\author{Ram\'on H. Mart\'inez-Mayorquin $^1$, Anders Lansner $^1$ and Pawel Herman $^1$
 %
@@
 % Addresses and institutions (remove "1- " in case of a single institution)
-1- KTH - Dept of First Author \\
-Address of First Author's school - Country of First Author's
-school
+1- KTH Royal Institute of Technology \\
+Department of Computational Science and Technology - Stockholm, Sweden.
 %
@@
 \vspace{.1cm}\\ 
-2- KTH - Dept of Second Author \\
-Address of Second Author's school - Country of Second Author's school 
-%
-\vspace{.1cm}\\
-3 School of Second Author - Dept of Second Author \\
-Address of Second Author's school - Country of Second Author's school\\ 
 }
@@
 
-\begin{abstract}
-Type your 100 words abstract here. Please do not modify the style
-of the paper. In particular, keep the text offsets to zero when
-possible (see above in this `ESANNV2.tex' sample file). You may
-\emph{slightly} modify it when necessary, but strictly respecting
-the margin requirements is mandatory (see the instructions to
-authors for more details).
+\begin{abstract} A rate-based recurrent network model equipped with a basic Hebbian learning rule including synaptic gating and memory traces is shown in this study to effectively learn sequences of activity patterns.
+The proposed computational framework facilitates understanding key mechanisms and the role of parameters in controlling the network performance and fundamental dynamics.
+We demonstrate among others how the sequence recall speed can be flexibly adjusted.
 \end{abstract}
@@
 \section{Introduction}
-
-This is a sample file. Please use this file to correctly typeset a
-submission to the ESANN conference. The associated pdf file will
-help you to have an idea of what your paper should look like.
+There is a growing interest in studying the capability to learn, store and flexibly recall sequential patterns in neural network models.
+Processing sequential information bears particular relevance to a wide scope of cognitive functionality and behavior in humans \cite{lashley1951problem}.
+Hebb proposed that sequential activation of neural cell assemblies, referred to as phases sequences, form the basis of the human thought process \cite{hebb2005organization}.
+The question of how a novel sequence of items is stored and recalled in the corrected order has been extensively studied in the context of memory systems among others.
+Besides intensive efforts in this direction undertaken in cognitive psychology \cite{hurlstone2014memory} the problem of sequence learning has also attracted a lot of theoretical interest.
+It has resulted in the development of different computational network models simulating the brain's capabilities to process sequence information.
+Among them are network approaches developed in the realm of rate based models that offer a generic context for sequence memory function \cite{amari1972learning}\cite{kohonen1977principle}\cite{sandberg2002bayesian}.
+The work presented here belongs to that category of studies and conceptually builds on \cite{sandberg2002bayesian}.
+In particular we rely on a simple recurrent neural network architecture and incorporate the concept of synaptic filtering into a basic Hebbian learning rule with synaptic gating \cite{andrew2003spiking} to study the network's capabilities to learn and recall sequences of memory items.
+The proposed rate based model constitutes a well-controlled framework for studying the fundamental synaptic mechanisms and identifying key parameters responsible for sequence processing capabilities in a dynamical connectionist system with all-to-all recurrent connectivity pattern.
+%
+%A prominent share of this computaitonal endevaour is linked to spiking neural network approaches to learn precise patterns of spike sequences with biologically plausible local learning rules \cite{ans1994neural}\cite{dehaene1987neural}\cite{gutig2006tempotron}\cite{ponulak2010supervised}.
 
 \section{Methods}
-The neural network is composed of firing rate units with the following properties.
-First the individual units receive current from all the other units, that is, the system is a recurrently connected neural network with all-to-all connectivity.
-All the neurons send current to each other and the accumulated current is then  thresholded $\theta$ by the logistic function $\Phi(x) = \frac{1}{1 + \exp(-Gx)}$.
-This whole process is condensed into equation \ref{eq1}.
-The current then is integrated with a time constant $\tau_m$, the integration time of the unit, to determine the activation of the unit.
-In order to account for the temporal effects we introduce the idea of z-filters that keep information about the activity during a time period $\tau_z$.
-We can see in equation \ref{eq2} that the filter z follows the activity of the unit at any point in time.
+
+The recurrent neural network in Fig.
+\ref{Fig:diagrams} is composed of firing rate units with the input-output characteristics described by the logistic function $\Phi(x) = \frac{1}{1 + \exp(-Gx)}$ with gain G.
+The synaptic inputs are aggregated across the entire network through all-to-all connections weighted by synaptic weights $W$, and the activation of unit $U_i$, is then consistently determined by integrating synaptic inputs with the time constant $\tau_{m}$ (\ref{eq1}).
+To account for the temporal effect of synaptic smoothing we introduce the idea of z-filters  \cite{tully2016spike} that maintain the exponentially decaying traces of the activity with the time constant of $\tau_z$ (\ref{eq2}).
+This allows for keeping synaptic memory of the unit activity and influencing other post-synaptic units in the network even when the original pre-synaptic unit is not longer active.
 
@@
 
-This system allows for the units to keep influencing the other units through $W$ even when they are not active anymore.
-We will see how this is fundamental for the sequence processing capabilities of our system.
+In this work we examine two types of learning rules, pre-synaptic (3) and post-synaptic (4)
+gating \cite{andrew2003spiking}, which equip the proposed network with the capability to learn
+how to recall sequential activity patterns.
 
-The connectivity of the system is as depicted in figure \ref{Fig:diagram}.
-Note that that from each unit emanate two excitatory connections, one that self-reinforces the unit activity $Exc_{self}$ and another that excites the following neuron (contributing to the sequential structure of the activity) $Exc_T$.
-At the same time every unit inhibits all the units except the next one with $Inh$.
-Of special relevance are the backwards inhibitory projections that suppress the unit once the next unit in the sequence is activated
+\begin{align}
+\tag{3 pre-synaptic}
+\tau_w \dfrac{dw}{dt} &= (w_{max} - w) z_{pre}z_{post} + (w_{min} - w) (1 - z_{post})z_{pre} \label{eq:pre-synaptic rule}  \\
+\tag{4 post-synaptic}
+\tau_w \dfrac{dw}{dt} &= \underbrace{(w_{max} - w)}_{\text{saturation}} \underbrace{z_{pre}z_{post}}_{\text{Hebbian}} + \underbrace{(w_{min} - w)}_{\text{saturation}}\underbrace{ (1 - z_{pre})z_{post}}_{\text{synaptic-gating}} \label{eq:post-synaptic rule}
+\end{align}
 
-\begin{figure}[h!]
+\begin{figure}[H]
 \centering
-\includegraphics[scale=1.5]{diagram.eps}
-\caption{In this diagram we can appreciate the overall connectivity of the system.
-Note that the inhibitory connections are only fully depicted for the first unit for the sake of clarity.}\label{Fig:diagram}
+\includegraphics[scale=0.8]{diagram_mixed.eps}
+\caption{a) Scheme of the overall network connectivity of the system.
+Note that the inhibitory connections are only fully depicted for the first unit for the sake of clarity.
+b) Conceptual illustration of the mechanism of learning induced synaptic changes.
+Here the a black filling stands for an active unit.
+} \label{Fig:diagrams}
 \end{figure}
+
+The first term in (\ref{eq:pre-synaptic rule}) and (\ref{eq:post-synaptic rule}) is an associative Hebbian contribution \cite{hebb2005organization} and its role is to strengthen the connections between activated units.
+Importantly however, in comparison with the original formulation \cite{andrew2003spiking}, we use the z-filters instead of the unit activities.
+This provides the scope for making associative connections between units even when either of the two units is no longer active provided that its memory trace is retainedby the synapse.
+As mentioned earlier, the synaptic trace decays exponentially with the time constant of $\tau_z$ (\ref{eq2}), and to preserve serial recall order we introduce asymmetrical time constants for z-pre and z-post : $\tau_{z_{pre}} > \tau_{z_{post}}$.
+The second term is called synaptic gating \cite{andrew2003spiking}, which facilitates the formation of inhibitory connections whenever the activity (or its z-filtered trace) of one unit is not accompanied with the activity of the other unit.
+We examine two variations of this term to explore the functional implications of both pre- and post-synaptic gating.
+Finally, the weights are kept between $w_{min}$ and $w_{max}$ limits.
+The nature of synaptic changes under the synaptic gating rule is illustrated in Fig.
+\ref{Fig:diagrams}.
+
+%\begin{figure}[h!]
+%\centering
+%\includegraphics[scale=1.3]{plasticity_diagram.eps}
+%\caption{Conceptual illustration of the mechanism of learning induced synaptic changes.
+Here the a black filling stands for an active unit a) Hebbian associative plasticity that implies a weight increase when both pre- and pos-synaptic units are simultaneously active b) In the case of the post-synaptic rule the weight is decreased if the post-synaptic unit itself active without the pre-synaptic unit being active.
+c) In the case of the pre-synaptic rule the weight is decreased if the pre-synaptic unit is itself active without the post-synaptic unit being active.
+d) If neither pre- nor post-synaptic units are active, there is no weight modification.}\label{Fig:plasticity diagram}
+%\end{figure}
+%
+As a result of learning described above, three types of connections emerge from each unit (Fig.
+\ref{Fig:diagrams}): two excitatory connections - self-reinforcing unit activity, $Exc_{self}$, and exciting the following unit in the sequence (contributing to the sequential structure of the activity) $Exc_T$, and a range of inhibitory connections, $Inh$, to all the remaining units (among inhibitory connections of special relevance are the backwards inhibitory projections that suppress the unit activity once the next unit in the sequence gets activated).
 
@@
 
-In order to recall a particular sequence we cue the first element.
-This means that we artificially activate the first unit (clamping) of the sequence for a short time.
-After this initial cue, if the weights are set right, the neural network continues evolving on its own and a particular sequence is recalled.
-In figure \ref{Fig:recall} we can appreciate a successful recall of a sequence.
+To recall a particular sequence we cue the first sequence item (unit) by injecting a short input lasting around $\tau_z$.
+After this initial cue, if the weights have the right configuration as a result of training, the neural network continues to visit subsequent items in the learned sequence to the very
+end, as shown in Fig. \ref{Fig:recall}.
 
@@
 d) The connectivity matrix $W$.
-The weights of the matrix can be divided on the excitatory connections: the ones that induce transitions $Exc_{T}$ and the ones that stabilize the unit $Exc_{self}$ and the inhibitory connections $Inh$.}\label{Fig:recall}
+The weights of the matrix $W$ can be divided into the excitatory connections: the ones on the diagonal that stabilize the unit $Exc_{self}$, the ones under the diagonal that induce transitions $Exc_{T}$ and finally the ones that stabilize the unit  and the inhibitory connections $Inh$ which dominate the matrix $W$.
+We use different colors to represent different items.
+}\label{Fig:recall}
 \end{figure}
 
-%It is important to note that the weights of the connectivity matrix determine the particular behavior that the network will develop.
-In the diagonal of the connectivity matrix we have the weights of the excitation connections that keep the unit active once it already became activited $Exc_{self}$.
-That is, they sustain a positive feed-back loop that keeps the unit active.
-Below the diagonal, on the other hand, we have the weights of the connections that make the sequence transition from one state to the next $Exc_{T}$.
-Finally, the rest of the matrix consists of the weights for the inhibitory connections.
-Of special significance are the backward inhibitory projections above the diagonal, their role is to suppress the last pattern of the sequence once the next unit is activated.
-
-It is now instructive to detail at the level of the dynamics how the sequential activity is propagated.
-After the first unit $x_1$ has been activated for long enough due to clamping the filter $z_1$ will start catching up with the activity by increasing its value.
-When this happens the unit corresponding to $x_2$ will receive a growing current with value $Exc_{T} \times z_1$.
-As this process continues there will a point in time (exactly at $Exc_{T}(1 - e^{-\frac{t}{\tau_z}}) - \theta = 0$) when $x_2$ will receive enough current to become activated.
-This, will again will prompt the filter $z_2$ to start catching up initializing the process described again.
-On the other hand as soon as the second filter starts growing $z_2$ the first unit will start receiving an inhibitory current $Inh \times z_2$ due to the backward inhibitory projection, this will in turn suppress the unit $x_1$ allowing the sequence to propagate.
-Finally, once the unit $x_1$ is finally suppressed the excitatory current mediated by $Exc_T$ will be diminished which in most cases would also suppress the activity of $x_2$ if were not for the self-sustaining current $Exc_{self} \times z_2$.
-In short, the term $Exc_T$ ensures the transition of the activity from one unit of the sequence to the next whereas the competition between $Exc_{self}$ and $Inh$ first lock the pattern unit and then finally suppress it allowing the sequence to evolve towards its end without hindrance.
+The dynamics of the sequential activity propagation can be easily explained.
+After the first unit $U_1$ has been activated by the external input, the corresponding filter $z_1$ starts following the activity $x_i$.
+In consequence, the subsequent unit $U_2$ receives a growing current with value $z_1 Exc_T$ because of the forward connectivity.
+This process continues until the point in time when $Exc_{T}(1 - e^{-\frac{t}{\tau_z}}) - \theta = 0$ implying that $U_2$ receives enough current to become activated.
+This in turn prompts the filter $z_2$ to follow the activity $x_2$ initializing again the process described above.
+On the other hand as soon as the synaptic trace corresponding to $x_2$ starts growing, the preceding unit $U_1$ beings to receive an inhibitory current $z_2 Inh$ due to the backward inhibitory projection.
+This in turn suppresses the activity of unit $U_1$ facilitating the propagation of the sequence.
+Finally once the unit $U_1$ is completely suppressed the excitatory current mediated by $Exc_T$ is diminished, which in most cases would also suppress the activity $x_2$ if it was not for the self-sustaining current $z_2 Exc_{self}$.
+In short, the term $Exc_T$ ensures the transition of the activity from one network state (sequence item) to another one (the next item in the sequence) whereas the competition between $Exc_{self}$ ad $Inh$ first lock the pattern unit and then finally suppresses it enabling the sequence to continue its trajectory.
 
 With the knowledge above in mind we can envision the following  failure scenarios for the sequential dynamics.
-First, if the magnitude of $Exc_{self}$ the activity will not be big enough to overcome the threshold and the sequence will not propagate.
-Second, if $Exc_{Self}$ is enough to keep the unit active but $Exc_{T}$ is not big enough the next unit in the sequence will not be activated and the sequence will not propagate.
-If both $Exc_{self}$ is big enough to sustain the activity and $Exc_{T}$ is big enough to propagate it but the inhibitory connection $Inh$ is not enough to suppress the units the result will be that all the units in the sequence will be activated at the end.
+First, if the magnitude of $Exc_{self}$ weight is small the self-excitatory current is not sufficiently strong to overcome the threshold and the sequence stalls.
+Second, if $Exc_{self}$ is large enough to keep the unit active but the $Exc_{T}$ weight is weak, the next unit in the sequence does not become activated and the sequence remains stuck forever in the same state corresponding to the first pattern.
+Finally if both $Exc_{self}$ and $Exc_T$ are strong enough to sustain and propagate the activity respectively but the inhibitory weight $Inh$ is not sufficiently large to suppress the activity, all the units in the sequence become co-activated at the end.
 
@@
 
-A common problem in sequence learning is to scale the time duration of the sequence according to the needs of the required context.
-In terms of a neural network it means that once the dynamics of sequence recalling are stored in the shape of the weights we can control the \textit{recall time} of the sequence with the internal parameters of the network where the recall time is the amount of time that a particular unit is active during a sequence recall.
-In our network there are a two parameters that a priori are strong candidates to regulate the recall time: the transition weight $Exc_T$ by controlling how fast you activate the next unit in the sequence and the threshold $\theta$ by controlling how long it takes for the current to activate the next unit in a sequence.
-Here we show the exact results for $Exc_T$ but the reasoning for $\theta$ is analogous.
+A common problem in sequence learning is to scale the time duration of the sequence according to the contextual demands \cite{murray2017learning}.
+For a neural network it means that the dynamics of sequence recall can be flexibly controlled with the network parameters after the training process is completed.
+The recall time can be defined as the temporal length of the activation of the individual units during the sequence retrieval.
+In the network there are two key parameters that are directly implicated in the regulation of the recall time: the transition weight, $Exc_T$ and the threshold, $\theta$, which control how long it takes to recruit the next unit in a sequence.
+By taking into account the competition between all the currents involved and the threshold we can calculate exactly the point at which the unit will become suppressed:
 
-In order to calculate the recall time we make the following consideration.
-Once the unit $x_1$ becomes activated the filter $z_1$ starts increasing its value and the unit $x_2$ will be receiving an excitatory current $Exc_T \times z_1$.
-The point in time t at which the this current overcomes the threshold will be the recall time.
-Using that reasoning we can solve the following equation to calculate the recall time: $Exc_{T}(1 - e^{-\frac{t}{\tau_z}}) - \theta = 0$.
-If we do we end up with the following expression: $T_{recall} = \tau_z \log(\frac{Exc_T}{Exc_T - \theta})$.
-This expression tells us that we actually have a singularity at $Exc_T = \theta$, which for our purposes means that can make the recall time arbitrarily by setting this two parameters close enough.
-In our neural network however, there are more currents into play because of the all-to-all connectivity, we can include those corrections and we end up with an expression of the form $T_{recall} = \tau_z \log(\frac{Exc_T + Inh}{Exc_T - \theta})$ which also possess the after-mentioned property.
-We can see an illustration of the wide dynamical range of our system in figure \ref{Fig:dynamical_range} where we also show two examples of sequence recall with different $Exc_T$ values and therefore with different recall times.
+\begin{align}
+T_{recall} = \tau_z \log\Big(\frac{Exc_T + Inh}{Exc_T - \theta}\Big) \label{eq:recall_complex}
+\end{align}
+
+This expression has a singularity at $Exc_T = \theta$, which implies that the recall time can be made arbitrarily long by setting $Exc_T$ close to the threshold.
+Fig.
+\ref{Fig:dynamical_range} illustrates the wide dynamical range of the recall time in our network model.
 
@@
 \centering
-\includegraphics[scale=0.25]{dynamical_range.eps}
+\includegraphics[scale=0.20]{dynamical_range.eps}
 \caption{The network posses a wide dynamical range.
@@
 \subsection{Learning Rule}
+Our learning protocol consists in fixing (clamping) each unit of the sequence in succession for a given amount of time (training time).
+We perform this process an epochs number of times.
+It is important to note that while learning there is a period of one second between each epoch or presentation that serves the purposes of avoiding self-interference.
 
-\begin{align}
-\tau_w \dfrac{dw}{dt} &= (w_{max} - w) z_{pre}z_{post} + (w_{min} - w) (z_{post} - 1)z_{pre} 
-\end{align}
+We illustrate the results of training in Fig.
+\ref{Fig:epochs} with two connectivity matrix capable of recalling, note the effects of $\tau_z$ in the strength and extent of the $Exc_T$ effects.
+Furthermore we can appreciate that the weights converge and stabilized with few epochs of training.
 
-\begin{figure}[h!]
+\begin{figure}[H]
 \centering
-\includegraphics[scale=1.3]{plasticity_diagram.eps}
-\caption{This is how plasticity works}\label{Fig:plasticity diagram}
+\includegraphics[scale=0.25]{training_rule.eps}
+\caption{The effect of training with the learning rule.
+a) An example of training with a small value of $tau_z=0.050s$.
+b) Same procedure as in a but with a longer time constant $\tau_z=0.150s$.
+c) evolution in time of the connection weights from the first to the second unit $Exc_T$, the one from the second back to the first unit $Inh$ and finally of the second unit into itself $Exc_{self}$.
+Five epochs of length $2s$ are shown here, note that this is enough time for the weights to stabilize.}\label{Fig:epochs}
 \end{figure}
 
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=0.3]{training_rule.eps}
-\caption{The training in action}\label{Fig:epochs}
-\end{figure}
+%\begin{figure}[H]
+%\centering
+%\includegraphics[scale=0.3]{training_rule_quantities_pre_rule_True.eps}
+%\caption{Learning dynamics of the pre-synaptic learning rule.}\label{Fig:learning_quantities_pre}
+%\end{figure}
+%
+%
+%\textbf{Training time}
+%\begin{itemize}
+%\item  $Exc_{self}$: The longer the training time the longer the filter is activated with itself and therefore it grows.
+But given that the pre and post filters do not coincide completely in the temporal domain there are negative contributions which prevent this from reaching the saturation value $w_{max}$
+%\item  $Inh$: In this case the the negative contributions come from the time that the units are not activated together with their counterparts which is most of the time (except in the brief transition period) and therefore the value saturates quickly to $w_{min}$.
+%\item $Exc_{T}$: In this case what matters is the transition time between two 
+%\end{itemize} 
+%
+%
+%\begin{figure}[H]
+%\centering
+%\includegraphics[scale=0.3]{training_rule_quantities_pre_rule_False.eps}
+%\caption{Learning dynamics of the post-synaptic learning rule}\label{Fig:learning_quantities_post}
+%\end{figure}
 
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=0.3]{training_rule_quantities.eps}
-\caption{The dynamics of the learning rule}\label{Fig:learning_quantities}
-\end{figure}
-
-\textbf{Training time}
-\begin{itemize}
-\item  $Exc_{self}$: The longer the training time the longer the filter is activated with itself and therefore it grows.
-But given that the pre and post filters do not coincide completely in the temporal domain there are negative contributions which prevent this from reaching the saturation value $w_{max}$
-\item  $Inh$: In this case the the negative contributions come from the time that the units are not activated together with their counterparts which is most of the time (except in the brief transition period) and therefore the value saturates quickly to $w_{min}$.
-\item $Exc_{T}$: In this case what matters is the transition time between two 
-\end{itemize}
-
-\subsection{Disambiguation}
-
-\section{Discussion}
-
-\section{Citation}
-This ESANNV2.tex file defines how to insert references, both for
-BiBTeX and non-BiBTeX users.  Please read the instructions in this
-file.
+%\section{Discussion}
 

### 2019-01-18 play.py

@@ -27,7 +27,7 @@ tau_m = 0.010
 I_cue = 0
 I_end = 5
 T = 10.0
-pre_rule = False
+pre_rule = True
 
 
 pattern = 3
@@ -64,11 +64,9 @@ if not saved_matrix:
 
         duration = get_recall_duration_for_pattern(x_history, pattern, dt)
 
-
         self = w[pattern, pattern]
         exc = w[pattern, pattern_from]
         inh = w[pattern_from, pattern]
-
 
         y[index, 0] = self
         y[index, 1] = exc
@@ -86,28 +84,3 @@ else:
     npz_file = np.load(outfile)
     X = npz_file['x']
     y = npz_file['y']
-
-# Here we do the classifier
-X_train, X_test, y_train, y_test = train_test_split(X, y, test_size=0.1, random_state=0)
-
-reg = linear_model.LinearRegression(fit_intercept=False, n_jobs=-1)
-reg.fit(X_train, y_train)
-score = reg.score(X_test, y_test)
-A = reg.coef_
-B = np.linalg.inv(A)
-print('score', score)
-
-
-# I want to calculate the values for which the self-excitation and the transition are positive
-positive_A = y[:, 0] > 0
-positive_T = y[:, 1] > 0
-negative_I = y[:, 2] < 0
-
-# Now the values that generated the valures in the correct domain are
-domain_indexes = positive_A * positive_T * negative_I
-X_right = X[domain_indexes, : ]
-
-
-
-
-
